---
title: "Manuscript 2 from Fernanda Hansen P. de Moraes Thesis - Cortical folding alterations in humans due to aging and diseases"
author: "Fernanda Hansen Pacheco de Moraes"
date: "18 jan 2022"
output:
  html_document: 
    fig_caption: yes
    fig_width: 8
    number_sections: yes
    theme: paper
    toc: yes
editor_options: 
  chunk_output_type: inline
---

Description of the procedures and analysis present in Manuscript 2,
**Establishing a baseline for human cortical folding morphological variables: a multicenter study**, at the Doctorate Thesis presented
to the Programa de Pós-Graduação em Ciências Médicas at the Instituto
D'Or de Pesquisa e Ensino as a partial requirement to obtain the
Doctorate Degree.

Part of the data used here cannot be shared due to restrictions of the
Ethic Committee. Data can be shared upon reasonable request to the
corresponding author. To fulfill these limitation, we will generate
random data to simulate the results.

Get in touch with us
([fernandahmoraes\@gmail.com](mailto:fernandahmoraes@gmail.com){.email})
in case any help is needed, our aim is to improve the code as needed!

# RMARKDOWN AND R SETUP
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = FALSE,
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)
```

```{r working directory}
setwd("~/GitHub/Typical-values")
```

```{r functions, message=FALSE, warning=FALSE}
## define functions

# test angular coeficinet versus theoretical value
test_coef <- function(reg, coefnum, val){
  co <- coef(summary(reg))
  tstat <- (co[coefnum,1] - val)/co[coefnum,2]
  2 * pt(abs(tstat), reg$df.residual, lower.tail = FALSE)
}

# wrap text
wrapper <- function(x, ...) paste(strwrap(x, ...), collapse = "\n")
```

```{r call packages}
library(readr)
library(tidyverse)
library(lubridate)
library(ggpubr)
library(kableExtra)
library(broom)
# library(MASS)
library(lme4)
library(lsmeans)


```

```{r} 
# COLOR BLIND PALETTE WITH BLACK
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cbbPalette2 <- c("#D55E00", "#E69F00", "#56B4E9", "#0072B2", "#CC79A7", "#009E73", "#F0E442")
```

# set seed for random process

```{r}
set.seed(1)
```

```{r import files}
dados_datasetscomp <- read_csv("dados_datasetscomp.csv")

dados_datasetscomp <- dados_datasetscomp %>%
  filter(Sample != "ds000030",
         Diagnostic != "MCI")

write.csv(dados_datasetscomp,"dados_datasetscomp2.csv")
```

# DATA PREPARATION
```{r create new variables}
# estimate cortical folding variables
dados_datasetscomp <- dados_datasetscomp %>%
  mutate(
    # create new variables
    logAvgThickness = log10(AvgThickness),
    logTotalArea = log10(TotalArea),
    logExposedArea = log10(ExposedArea),
    localGI = TotalArea / ExposedArea,
    k = sqrt(AvgThickness) * TotalArea / (ExposedArea ^ 1.25),
    K = 1 / 4 * log10(AvgThickness ^ 2)  + log10(TotalArea) - 5 / 4 * log10(ExposedArea),
    S = 3 / 2 * log10(TotalArea) + 3 / 4 * log10(ExposedArea) - 9 /  4 * log10(AvgThickness ^
                                                                                 2) ,
    I = log10(TotalArea) + log10(ExposedArea) + log10(AvgThickness ^ 2),
    c = as.double(ifelse(
      ROI == "hemisphere", NA, 4 * pi / GaussianCurvature
    )),
    TotalArea_corrected = ifelse(ROI == "hemisphere", TotalArea, TotalArea * c),
    ExposedArea_corrected = ifelse(ROI == "hemisphere", ExposedArea, ExposedArea * c),
    logTotalArea_corrected = log10(TotalArea_corrected),
    logExposedArea_corrected = log10(ExposedArea_corrected),
    localGI_corrected = ifelse(
      ROI == "hemisphere",
      TotalArea / ExposedArea,
      TotalArea_corrected / ExposedArea_corrected
    ),
    k_corrected = ifelse(
      ROI == "hemisphere",
      sqrt(AvgThickness) * log10(TotalArea) / (log10(ExposedArea) ^ 1.25),
      sqrt(AvgThickness) * log10(TotalArea_corrected) / (log10(ExposedArea_corrected ^
                                                                 1.25))
    ),
    K_corrected =  ifelse(
      ROI == "hemisphere",
      1 / 4 * log10(AvgThickness ^ 2) + log10(TotalArea) - 5 / 4 * log10(ExposedArea),
      1 / 4 * log10(AvgThickness ^ 2) + log10(TotalArea_corrected) - 5 / 4 * log10(ExposedArea_corrected)
    ),
    I_corrected = ifelse(
      ROI == "hemisphere",
      log10(TotalArea) + log10(ExposedArea) + log10(AvgThickness ^ 2) ,
      log10(TotalArea_corrected) + log10(ExposedArea_corrected) + log10(AvgThickness ^ 2)
    ),
    S_corrected = ifelse(
      ROI == "hemisphere",
      3 / 2 * log10(TotalArea) + 3 / 4 * log10(ExposedArea) - 9 /  4 * log10(AvgThickness ^ 2) ,
      3 / 2 * log10(TotalArea_corrected) + 3 / 4 * log10(ExposedArea_corrected) - 9 /  4 * log10(AvgThickness ^ 2)
    ),
    Knorm = K_corrected / sqrt(1 + (1 / 4) ^ 2 + (5 / 4) ^ 2),
    Snorm = S_corrected / sqrt((3 / 2) ^ 2 + (3 / 4) ^ 2 + (9 / 4) ^ 2),
    Inorm = I_corrected / sqrt(1 ^ 2 + 1 ^ 2 + 1 ^ 1)
  )

# create age intervals
dados_datasetscomp$Age_interval <- cut(dados_datasetscomp$Age,
                                       breaks = c(0, 5, 10, 15, 20, 25, 30, 35, 40, 45, 50, 55, 60, 65, 70, 75, 80, 85, 90, 95, 100),
                                       right = FALSE,
                                       include.lowest = TRUE)

dados_datasetscomp$Age_interval10 <- cut(dados_datasetscomp$Age,
                                         breaks = c(0, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100),
                                         right = FALSE,
                                         include.lowest = TRUE)
```

```{r data cleanup}
dados_all <- dados_datasetscomp %>% filter(!is.na(logAvgThickness), ExposedArea != 0 | !is.na(localGI), !is.infinite(logExposedArea)) %>% 
  droplevels()

dados_datasetscomp <- dados_all

# dados_datasetscomp$Diagnostic <- relevel(dados_datasetscomp$Diagnostic, ref = "CTL")

```

# Deaging

```{r deaging}
# define age for deaging
Age.cor = 25

# DEAGING + HARMONIZATION FULL DATA ----

dados_datasetscomp_rate <-
  filter(dados_datasetscomp, Diagnostic == "CTL")

dados_datasetscomp_rate$Sample <-
  as.factor(dados_datasetscomp_rate$Sample)

dados_datasetscomp_rate$ROI <-
  factor(dados_datasetscomp_rate$ROI,
         levels = c("hemisphere", "F", "O", "P", "T"))

## AvgThickness ---
m.1 <-
  lme4::lmer(AvgThickness ~ Age * ROI + (1|Sample:ROI), data = dados_datasetscomp_rate)

re <- as_tibble(ranef(m.1)) %>%
  filter(grpvar == "Sample:ROI") %>%
  mutate(
    T_shift = condval,
    sd_T_shift = condsd,
    Sample = str_split(grp, pattern = ":", simplify = TRUE)[, 1],
    ROI = str_split(grp, pattern = ":", simplify = TRUE)[, 2]
  ) %>%
  dplyr::select(-c(condval, grpvar, term, condsd, grp))

Age.trend <- as_tibble(lstrends(m.1, ~ ROI, var = "Age")) %>%
  mutate(Age.trend_T = Age.trend) %>%
  dplyr::select(c(ROI, Age.trend_T))

dados_datasetscomp <- full_join(dados_datasetscomp, Age.trend) %>%
  full_join(re) %>%
  mutate(
    AvgThickness_shiftc = AvgThickness - T_shift,
    AvgThickness_age_decay = AvgThickness - Age.trend_T * (Age - Age.cor),
    AvgThickness_age_decay_shiftc = AvgThickness - T_shift - Age.trend_T *
      (Age - Age.cor),
    logAvgThickness_shiftc = log10(AvgThickness_shiftc),
    logAvgThickness_age_decay = log10(AvgThickness_age_decay),
    logAvgThickness_age_decay_shiftc = log10(AvgThickness_age_decay_shiftc)
  )

## TotalArea ---
m.1 <-
  lme4::lmer(TotalArea_corrected ~ Age * ROI + (1 | Sample:ROI), data = dados_datasetscomp_rate)

re <- as_tibble(ranef(m.1)) %>%
  filter(grpvar == "Sample:ROI") %>%
  mutate(
    AT_shift = condval,
    sd_AT_shift = condsd,
    Sample = str_split(grp, pattern = ":", simplify = TRUE)[, 1],
    ROI = str_split(grp, pattern = ":", simplify = TRUE)[, 2]
  ) %>%
  dplyr::select(-c(condval, grpvar, term, condsd, grp))

Age.trend <-
  as_tibble(lstrends(m.1, ~ ROI, var = "Age")) %>%
  mutate(Age.trend_AT = Age.trend) %>%
  dplyr::select(c(ROI, Age.trend_AT))

dados_datasetscomp <- full_join(dados_datasetscomp, re) %>%
  full_join(Age.trend) %>%
  mutate(
    TotalArea_shiftc = TotalArea_corrected - AT_shift,
    TotalArea_age_decay = TotalArea_corrected - Age.trend_AT * (Age - Age.cor),
    TotalArea_age_decay_shiftc = TotalArea_corrected - AT_shift - Age.trend_AT *
      (Age - Age.cor),
    logTotalArea_shiftc = log10(TotalArea_shiftc),
    logTotalArea_age_decay = log10(TotalArea_age_decay),
    logTotalArea_age_decay_shiftc = log10(TotalArea_age_decay_shiftc)
  )

## ExposedArea ---
m.1 <-
  lme4::lmer(
    ExposedArea_corrected  ~ Age * ROI + (1 | Sample:ROI), data = dados_datasetscomp_rate)

re <- as_tibble(ranef(m.1)) %>%
  filter(grpvar == "Sample:ROI") %>%
  mutate(
    AE_shift = condval,
    sd_AE_shift = condsd,
    Sample = str_split(grp, pattern = ":", simplify = TRUE)[, 1],
    ROI = str_split(grp, pattern = ":", simplify = TRUE)[, 2]
  ) %>%
  dplyr::select(-c(condval, grpvar, term, condsd, grp))

Age.trend <-
  as_tibble(lstrends(m.1, ~ ROI, var = "Age")) %>%
  mutate(Age.trend_AE = Age.trend) %>%
  dplyr::select(c(ROI, Age.trend_AE))

dados_datasetscomp <- full_join(dados_datasetscomp, re) %>%
  full_join(Age.trend) %>%
  mutate(
    ExposedArea_shiftc = ExposedArea_corrected - AE_shift,
    ExposedArea_age_decay = ExposedArea_corrected - Age.trend_AE * (Age - Age.cor),
    ExposedArea_age_decay_shiftc = ExposedArea_corrected - AE_shift - Age.trend_AE * (Age - Age.cor),
    logExposedArea_shiftc = log10(ExposedArea_shiftc),
    logExposedArea_age_decay = log10(ExposedArea_age_decay),
    logExposedArea_age_decay_shiftc = log10(ExposedArea_age_decay_shiftc)
  )

## AvgThickness ---
m.1 <-
  lme4::lmer(c ~ Age * ROI + (1|Sample:ROI), data = dados_datasetscomp_rate)

re <- as_tibble(ranef(m.1)) %>%
  filter(grpvar == "Sample:ROI") %>%
  mutate(
    c_shift = condval,
    sd_c_shift = condsd,
    Sample = str_split(grp, pattern = ":", simplify = TRUE)[, 1],
    ROI = str_split(grp, pattern = ":", simplify = TRUE)[, 2]
  ) %>%
  dplyr::select(-c(condval, grpvar, term, condsd, grp))

Age.trend <- as_tibble(lstrends(m.1, ~ ROI, var = "Age")) %>%
  mutate(Age.trend_c = Age.trend) %>%
  dplyr::select(c(ROI, Age.trend_c))

dados_datasetscomp <- full_join(dados_datasetscomp, Age.trend) %>%
  full_join(re) %>%
  mutate(
    c_shiftc = c - c_shift,
    c_age_decay = c - Age.trend_c * (Age - Age.cor),
    c_age_decay_shiftc = c - c_shift - Age.trend_c *
      (Age - Age.cor)
  )

# ----
dados_datasetscomp <- dados_datasetscomp %>%
  mutate(
    localGI_age_decay = TotalArea_age_decay/ExposedArea_age_decay,
    localGI_shiftc = TotalArea_shiftc/ExposedArea_shiftc,
    localGI_age_decay_shiftc = TotalArea_age_decay_shiftc/ExposedArea_age_decay_shiftc,
    K_age_decay = log10(TotalArea_age_decay) + 1/4*log10(AvgThickness_age_decay^2) - 5/4*log10(ExposedArea_age_decay),
    K_shiftc = log10(TotalArea_shiftc) + 1/4*log10(AvgThickness_shiftc^2) - 5/4*log10(ExposedArea_shiftc),
    K_age_decay_shiftc = log10(TotalArea_age_decay_shiftc) + 1/4*log10(AvgThickness_age_decay_shiftc^2) - 5/4*log10(ExposedArea_age_decay_shiftc),
    I_age_decay = log10(TotalArea_age_decay) + log10(ExposedArea_age_decay) + log10(AvgThickness_age_decay^2),
    I_shiftc = log10(TotalArea_shiftc) + log10(ExposedArea_shiftc) + log10(AvgThickness_shiftc^2),
    I_age_decay_shiftc = log10(TotalArea_age_decay_shiftc) + log10(ExposedArea_age_decay_shiftc) + log10(AvgThickness_age_decay_shiftc^2),
    S_age_decay = 3/2*log10(TotalArea_age_decay) + 3/4*log10(ExposedArea_age_decay) - 9/4*log10(AvgThickness_age_decay^2),
    S_shiftc = 3/2*log10(TotalArea_shiftc) + 3/4*log10(ExposedArea_shiftc) - 9/4*log10(AvgThickness_shiftc^2),
    S_age_decay_shiftc = 3/2*log10(TotalArea_age_decay_shiftc) + 3/4*log10(ExposedArea_age_decay_shiftc) - 9/4*log10(AvgThickness_age_decay_shiftc^2),
    Knorm_shiftc = K_shiftc / sqrt(1 + (1 / 4) ^ 2 + (5 / 2) ^ 2),
    Snorm_shiftc = S_shiftc / sqrt((3 / 2) ^ 2 + (3 / 4) ^ 2 + (9 / 4) ^ 2),
    Inorm_shiftc = I_shiftc / sqrt(1 ^ 2 + 1 ^ 2 + 1 ^ 2),
    Knorm_age_decay = K_age_decay / sqrt(1 + (1 / 4) ^ 2 + (5 / 2) ^ 2),
    Snorm_age_decay = S_age_decay / sqrt((3 / 2) ^ 2 + (3 / 4) ^ 2 + (9 / 4) ^ 2),
    Inorm_age_decay = I_age_decay / sqrt(1 ^ 2 + 1 ^ 2 + 1 ^ 2),
    Knorm_age_decay_shiftc = K_age_decay_shiftc / sqrt(1 + (1 / 4) ^ 2 + (5 / 2) ^ 2),
    Snorm_age_decay_shiftc = S_age_decay_shiftc / sqrt((3 / 2) ^ 2 +  (3 / 4) ^ 2 + (9 / 4) ^ 2),
    Inorm_age_decay_shiftc = I_age_decay_shiftc / sqrt(1 ^ 2 + 1 ^ 2 + 1 ^ 2)
  )


```

# RESULTS

## Data description

```{r, echo=FALSE, message=FALSE, warning=FALSE}
filter(dados_datasetscomp,
       Sample != "Mota&Houzel2015",
       ROI == "hemisphere") %>%
  group_by(Sample, Diagnostic) %>%
  summarise(
    N = n_distinct(SUBJ),
    age = paste(signif(mean(Age), 2), "±", signif(sd(Age), 2)),
    age_range = paste(signif(min(Age), 2), "; ", signif(max(Age), 2)),
    T =  paste(signif(mean(AvgThickness), 2), "±", signif(sd(AvgThickness), 2)),
    AT =  paste(signif(mean(TotalArea), 2), "±", signif(sd(TotalArea), 2)),
    AE =  paste(signif(mean(ExposedArea), 2), "±", signif(sd(ExposedArea), 2)),
    K =  paste(signif(mean(K), 2), "±", signif(sd(K), 2)),
    S =  paste(signif(mean(S), 2), "±", signif(sd(S), 2)),
    I =  paste(signif(mean(I), 2), "±", signif(sd(I), 2))
  ) %>%
  kable(digits = 2) %>%
  kable_styling()
```

## Correlation between K, S, and I with Age for the Healthy Control group

### K 

```{r}
figS6a <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"),
       aes(x = Age , y = K)) +
  geom_point(aes(color = Sample, fill = Sample, alpha = 0.4)) +
  geom_smooth(color = "black", method = "lm") +
  theme_pubr() +
  guides(alpha = "none")+ 
  labs(x = "Age [years]") +
  scale_x_continuous(limits = c(0,100))

cor.test(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL")$K, filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL")$Age)
```

### S

```{r}
figS6b <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"),
       aes(x = Age, y = S)) +
  geom_point(aes(color = Sample, fill = Sample, alpha = 0.4)) +
  geom_smooth(color = "black", method = "lm") +
  theme_pubr() +
  guides(alpha = "none", color = FALSE,fill = FALSE)+ 
  labs(x = "Age [years]") +
  theme(legend.position= "none") +
  scale_x_continuous(limits = c(0,100))

cor.test(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL")$S, filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL")$Age)
```

### I

```{r}
figS6c <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"),
       aes(x = Age, y = I)) + geom_point(aes(color = Sample, fill = Sample, alpha = 0.4)) + 
  geom_smooth(color = "black", method = "lm") + theme_pubr() +
  guides(alpha = "none", color = FALSE,fill = FALSE)+ 
  labs(x = "Age [years]") +
  theme(legend.position= "none") +
  scale_x_continuous(limits = c(0,100))

cor.test(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL")$I, filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL")$Age)

```

```{r}

figS6 <- ggarrange(figS6a, ggarrange(figS6b, figS6c,labels = c("B","C"), nrow = 1, ncol = 2, font.label = list(size = 11)),labels = c("A"), nrow = 2, ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")

ggsave("figS6.pdf", plot = figS6, dpi=1200, width = 18, height = 22, units = "cm", device = "pdf")


```

**FIGURE 1**
```{r figure1, fig.cap="\\label{fig:figure1}Through age, every Healthy Control subject for the independent morphological component, K, S, and I. For HCPr900 and AHEAD, ages are determined by an interval of years, thereby considering the mean age. The solid line represents a linear regression with a 95% confidence interval.", fig.height=8.66, fig.width=7.1}

figS6
```

## Slope

```{r}
summary(lm(
  1 / 2 * log10(AvgThickness) + log10(TotalArea) ~ log10(ExposedArea),
  data = filter(dados_datasetscomp,
       Sample != "Mota&Houzel2015",
       ROI == "hemisphere"),
  na.action = na.omit
))

## Displays confidence interval
# tidy(lm(
#   1 / 2 * log10(AvgThickness) + log10(TotalArea) ~ log10(ExposedArea),
#   data = dados_hemi_v1,
#   na.action = na.omit
# ), conf.int = TRUE)

paste(
  "Student's t test comapring slope with theoretical value 1.25. t = ",
  signif(abs(coef(summary(
    lm(
      1 / 2 * log10(AvgThickness) + log10(TotalArea) ~ log10(ExposedArea),
      data = filter(dados_datasetscomp,
       Sample != "Mota&Houzel2015",
       ROI == "hemisphere"),
      na.action = na.omit
    )
  ))[2, 1] - 1.25) / coef(summary(
    lm(
      1 / 2 * log10(AvgThickness) + log10(TotalArea) ~ log10(ExposedArea),
      data = filter(dados_datasetscomp,
       Sample != "Mota&Houzel2015",
       ROI == "hemisphere"),
      na.action = na.omit
    )
  ))[2, 2], 3) 
)
paste(
  "Student's t test comapring slope with theoretical value 1.25. p value = ",
  signif(test_coef(
    lm(
      1 / 2 * log10(AvgThickness) + log10(TotalArea) ~ log10(ExposedArea),
      data = filter(dados_datasetscomp,
       Sample != "Mota&Houzel2015",
       ROI == "hemisphere"),
      na.action = na.omit
    ),
    2,
    1.25
  ),3)
)
```

### Correlation 

```{r}
lm_Age <-
  filter(
    dados_datasetscomp,
    Sample != "Mota&Houzel2015",
    ROI == "hemisphere",
    Diagnostic == "CTL",
    Age_interval != "[0,5)",
    Age_interval != "[5,10)",
    Age_interval != "[10,15)"
  ) %>%
  group_by(Age_interval) %>%
  do(fit_Age = tidy(
    lm(
      1 / 2 * log10(AvgThickness) +  log10(TotalArea) ~  log10(ExposedArea),
      data = .,
      na.action = na.omit
    ),
    conf.int = TRUE
  )) %>%
  unnest(cols = c(fit_Age))

lm_Age <- lm_Age %>% mutate(Age_interval = as.double((str_sub(Age_interval,2,3))))

cor.test(filter(lm_Age, term == "log10(ExposedArea)")$estimate, filter(lm_Age, term == "log10(ExposedArea)")$Age_interval)

```

## Uncertainties estimation for baseline

## Multisite harmonization

## Baseline and rate estimates

### Slope

```{r}
lm_Age <-
  filter(
    dados_datasetscomp,
    Sample != "Mota&Houzel2015",
    ROI == "hemisphere",
    Diagnostic == "CTL",
    Age_interval != "[0,5)",
    Age_interval != "[5,10)",
    Age_interval != "[10,15)"
  ) %>%
  group_by(Age_interval) %>%
  do(fit_Age = tidy(
    lm(
      1 / 2 * log10(AvgThickness_shiftc) +  log10(TotalArea_shiftc) ~  log10(ExposedArea_shiftc),
      data = .,
      na.action = na.omit
    ),
    conf.int = TRUE
  )) %>%
  unnest(cols = c(fit_Age))

lm_Age <- lm_Age %>% mutate(Age_interval = as.double((str_sub(Age_interval,2,3))))

cor.test(filter(lm_Age, term == "log10(ExposedArea_shiftc)")$estimate, filter(lm_Age, term == "log10(ExposedArea_shiftc)")$Age_interval)

```

```{r cache=TRUE}
amostra_Coef <- filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL", Age_interval != "[0,5)", Age_interval != "[95,100]") %>%
    group_by(Age_interval) %>%
    do(fit_amostra = tidy(lm(1/2 * log10(AvgThickness_shiftc) + log10(TotalArea_shiftc) ~ log10(ExposedArea_shiftc), data = ., na.action = na.omit), conf.int = TRUE, conf.level = 0.95)) %>% 
    unnest(fit_amostra)

amostras_Coef_age <- filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL", Age_interval != "[0,5)", Age_interval != "[95,100]") %>%
    group_by(Age_interval) %>%
    summarise(
        N = n_distinct(SUBJ),
        min_age = min(Age),
        max_age = max(Age))

amostra_Coef <- full_join(amostra_Coef, amostras_Coef_age) %>% filter(term == "log10(ExposedArea_shiftc)")

fig_slope_ageinterval <- ggplot(amostra_Coef, aes(y = estimate, x = Age_interval)) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
    geom_point() +
  geom_line(group = 1) + 
  geom_hline(yintercept = 1.25, linetype = "dashed") +
    theme_pubr() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=9)
    ) + labs(y ="Slope (after harmonization)", x = "Age interval") +  geom_text(aes(label = N), nudge_y = 0.2)

ggsave("fig_slope_ageinterval.pdf", plot = fig_slope_ageinterval, dpi=1200, width = 18, height = 14, units = "cm", device = "pdf")
```

**FIGURE 3**
```{r figure3, fig.cap="\\label{fig:figure3}Slope alpha derived from the Cortical Folding model from Mota & Houzel through age for Healthy Controls. Data points from (0, 5] and [95, 100] years old were omitted due to the reduced data. The numbers on top of each point indicates the number of subjects at each interval. Mean value for all ages were 1.38+-0.012 and correlation of alpha and Age is Pearson's r~=~-0.69, p~=~0.0031 from [15,20) years old.", fig.height=5.5, fig.width=7.1}

fig_slope_ageinterval
```

## Diagnostic discrimination based on changing rates

