---
title: "Manuscript 2 from Fernanda Hansen P. de Moraes Thesis - Cortical folding alterations in humans due to aging and diseases"
author: "Fernanda Hansen Pacheco de Moraes"
date: "18 jan 2022"
output:
  html_document: 
    fig_caption: yes
    fig_width: 8
    number_sections: yes
    theme: paper
    toc: yes
editor_options: 
  chunk_output_type: inline
---

Description of the procedures and analysis present in Manuscript 2,
**Establishing a baseline for human cortical folding morphological variables: a multicenter study**, at the Doctorate Thesis presented
to the Programa de Pós-Graduação em Ciências Médicas at the Instituto
D'Or de Pesquisa e Ensino as a partial requirement to obtain the
Doctorate Degree.

Part of the data used here cannot be shared due to restrictions of the
Ethic Committee. Data can be shared upon reasonable request to the
corresponding author. To fulfill these limitation, we will generate
random data to simulate the results.

Get in touch with us
([fernandahmoraes\@gmail.com](mailto:fernandahmoraes@gmail.com){.email})
in case any help is needed, our aim is to improve the code as needed!

# RMARKDOWN AND R SETUP
```{r setup, include=FALSE}
knitr::opts_chunk$set(
	echo = TRUE,
	message = FALSE,
	warning = FALSE,
	cache = TRUE
)
```

```{r working directory}
setwd("~/GitHub/Typical-values")
```

```{r functions, message=FALSE, warning=FALSE}
## define functions

# test angular coeficinet versus theoretical value
test_coef <- function(reg, coefnum, val){
  co <- coef(summary(reg))
  tstat <- (co[coefnum,1] - val)/co[coefnum,2]
  2 * pt(abs(tstat), reg$df.residual, lower.tail = FALSE)
}

# wrap text
wrapper <- function(x, ...) paste(strwrap(x, ...), collapse = "\n")
```

```{r call packages}
library(readr)
library(tidyverse)
library(lubridate)
library(ggpubr)
library(kableExtra)
library(broom)
library(rstatix)

```

```{r} 
# COLOR BLIND PALETTE WITH BLACK
cbbPalette <- c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")
cbbPalette2 <- c("#CC79A7", "#E69F00", "#56B4E9", "#0072B2", "#CC79A7", "#009E73", "#F0E442")
```

# set seed for random process

```{r}
set.seed(1)
```

```{r import files}
dados <- read_csv("dados_AOMICID1000.csv")

```
# RESULTS

## Data description

```{r description, echo=FALSE}
filter(dados, ROI == "hemisphere") %>%
  group_by(Session) %>%
  summarise(
    N = n_distinct(SUBJ),
    age = paste(signif(mean(Age), 2), "±", signif(sd(Age), 2)),
    age_range = paste(signif(min(Age), 2), "; ", signif(max(Age), 2)),
    AvgT =  paste(signif(mean(AvgThickness), 2), "±", signif(sd(AvgThickness), 2)),
    AT =  paste(signif(mean(TotalArea), 2), "±", signif(sd(TotalArea), 2)),
    AE =  paste(signif(mean(ExposedArea), 2), "±", signif(sd(ExposedArea), 2)),
    K =  paste(signif(mean(K), 2), "±", signif(sd(K), 2)),
    S =  paste(signif(mean(S), 2), "±", signif(sd(S), 2)),
    I =  paste(signif(mean(I), 2), "±", signif(sd(I), 2))
) %>% kable(digits = 2) %>% kable_styling()

subj_1 <- unique(filter(dados, Session == 1)$SUBJ)
subj_2 <- unique(filter(dados, Session == 2)$SUBJ)
subj_3 <- unique(filter(dados, Session == 3)$SUBJ)

```

## Model application

```{r}
amostra_Coef <- filter(dados, ROI == "hemisphere") %>% group_by(Session) %>%
  do(fit_amostra = tidy(lm(1/2 * logAvgThickness + logTotalArea ~ logExposedArea, data = ., na.action = na.omit), conf.int = TRUE, conf.level = 0.95)) %>% 
  unnest(fit_amostra)

amostra_Coef %>%
  kable(digits = 2) %>%
  kable_styling()

```

```{r}
fig_A_1 <- ggplot(data = filter(dados, ROI == "hemisphere"), aes(log10(ExposedArea), 1 / 2 * log10(AvgThickness) + log10(TotalArea), color = as.factor(Session), fill = as.factor(Session))) +
  geom_smooth(method = "lm", se = FALSE) +
  geom_point() +
  geom_line(aes(group= interaction(SUBJ, hemi)), color = "black", alpha = 0.8) + 
  theme_pubr() +
  guides(shape = "none") +
  theme(
    axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    text = element_text(size = 10)
  )  +
  scale_fill_manual(values=cbbPalette2) +
  scale_colour_manual(values=cbbPalette2)

ggsave("fig_A_1.pdf", plot = fig_A_1, dpi=1200, width = 11, height = 11, units = "cm", device = "pdf")
```

```{r figure0, fig.cap="\\label{fig:figure0}Cortical folding model with the traced path for each hemisphere of each subject (gray line). The sample respects the model with slope alpha=1.23+-0.06, 95% confidence interval=(1.12,1.34).", fig.height=7.1, fig.width=7.1}
fig_A_1
```


## Sessions

```{r}

ses_T <- ggplot(filter(dados, ROI == "hemisphere"), aes(x = Session, y = AvgThickness)) +
    geom_boxplot()+
      geom_line(aes(group= interaction(SUBJ, hemi)), color = "grey", alpha = 0.8) + 
    theme_pubr() +
    theme(
        axis.title = element_text(size = 11),
        axis.text = element_text(size = 10),
        text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1, size=9)
    )  + labs(y = "AvgThickness [mm]", x = "Run")

# filter(dados, ROI == "hemisphere") %>%
#   group_by(Session) %>%
#   identify_outliers(AvgThickness)

res.aov <- anova_test(data = filter(dados, ROI == "hemisphere"), dv = AvgThickness, wid = c(SUBJ, hemi), within = Session)
get_anova_table(res.aov)

ses_K <- ggplot(filter(dados, ROI == "hemisphere"), aes(x = Session, y = K)) +
    geom_boxplot()+
      geom_line(aes(group= interaction(SUBJ, hemi)), color = "grey", alpha = 0.8) + 
    theme_pubr() +
    theme(
        axis.title = element_text(size = 11),
        axis.text = element_text(size = 10),
        text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1, size=9)
    )  + labs(y = "K", x = "Run")

res.aov <- anova_test(data = filter(dados, ROI == "hemisphere"), dv = K, wid = c(SUBJ, hemi), within = Session)
get_anova_table(res.aov)

ses_S <- ggplot(filter(dados, ROI == "hemisphere"), aes(x = Session, y = S)) +
    geom_boxplot()+
      geom_line(aes(group= interaction(SUBJ, hemi)), color = "grey", alpha = 0.8) + 
    theme_pubr() +
    theme(
        axis.title = element_text(size = 11),
        axis.text = element_text(size = 10),
        text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1, size=9)
    )  + labs(y = "S", x = "Run")

res.aov <- anova_test(data = filter(dados, ROI == "hemisphere"), dv = S, wid = c(SUBJ, hemi), within = Session)
get_anova_table(res.aov)

ses_I <- ggplot(filter(dados, ROI == "hemisphere"), aes(x = Session, y = I)) +
    geom_boxplot()+
      geom_line(aes(group= interaction(SUBJ, hemi)), color = "grey", alpha = 0.8) + 
    theme_pubr() +
    theme(
        axis.title = element_text(size = 11),
        axis.text = element_text(size = 10),
        text = element_text(size = 10),
        axis.text.x = element_text(angle = 45, hjust = 1, size=9)
    )  + labs(y = "I", x = "Run")

res.aov <- anova_test(data = filter(dados, ROI == "hemisphere"), dv = I, wid = c(SUBJ, hemi), within = Session)
get_anova_table(res.aov)

ses_var <- ggarrange(ses_T, ses_K, ses_S, ses_I, nrow = 2, ncol = 2, labels = c("A", "B", "C", "D"), font.label = list(size = 11))

ggsave("ses_var.pdf", plot = ses_var, dpi=1200, width = 18, height = 18, units = "cm", device = "pdf")

```

```{r figure1, fig.cap="\\label{fig:figure1}Trajectories for each hemisphere of each subject through the acquisition runs. There is no significant difference in means. (A) Average Cortical Thickness~[mm]: F(2, 198)=0.481, p=0.62, eta2[g]=0.000192; (B) K: F(2, 198)=0.138, p=0.87, eta2[g]=0.0000144; (C) S: F(2, 185)=0.352, p=0.69, eta2[g]=0.0000798; (D) I: F(2, 198)=0.775, p=0.46, eta2[g]=0.0000523.", fig.height=7.1, fig.width=7.1}
ses_var
```

## SUBJ variance

```{r}
SD <- filter(dados, ROI == "hemisphere") %>% group_by(SUBJ, hemi) %>%
  summarise(
  meanT = mean(AvgThickness),
  meanAT = mean(TotalArea),
  meanAE = mean(ExposedArea),
  meanGM = mean(GMvolume),
  meanK =  mean(K),
  meanS =  mean(S),
  meanI =  mean(I),
  meanGI = mean(localGI),
  sd_T = sd(AvgThickness),
  sd_AT = sd(TotalArea),
  sd_AE = sd(ExposedArea),
  sd_GM = sd(GMvolume),
  sd_K = sd(K),
  sd_S = sd(S),
  sd_I = sd(I),
  sd_GI = sd(localGI),
  SD_percent_T = sd_T*100/abs(meanT),
  SD_percent_AT = sd_AT*100/abs(meanAT),
  SD_percent_AE = sd_AE*100/abs(meanAE),
  SD_percent_GM = sd_GM*100/abs(meanGM),
  SD_percent_K = sd_K*100/abs(meanK),
  SD_percent_S = sd_S*100/abs(meanS),
  SD_percent_I = sd_I*100/abs(meanI),
  SD_percent_GI = sd_GI*100/abs(meanGI)
  )

SD %>%
  summarise(
  N = n_distinct(SUBJ),
  meanT = mean(meanT),
  meanAT = mean(meanAT),
  meanAE = mean(meanAE),
  meanGM = mean(meanGM),
  meanK =  mean(meanK),
  meanS =  mean(meanS),
  meanI =  mean(meanI),
  meanGI = mean(meanGI),
  sd_T = mean(sd_T),
  sd_AT = mean(sd_AT),
  sd_AE = mean(sd_AE),
  sd_GM = mean(sd_GM),
  sd_K = mean(sd_K),
  sd_S = mean(sd_S),
  sd_I = mean(sd_I),
  sd_GI = mean(sd_GI)
  ) %>%
  summarise(
  meanT = signif(mean(meanT),2),
  meanAT = signif(mean(meanAT),2),
  meanAE = signif(mean(meanAE),2),
  meanGM = signif(mean(meanGM),2),
  meanK =  signif(mean(meanK),2),
  meanS =  signif(mean(meanS),2),
  meanI =  signif(mean(meanI),2),
  meanGI = signif(mean(meanGI),2),
  sd_T = signif(mean(sd_T),2),
  sd_AT = signif(mean(sd_AT),2),
  sd_AE = signif(mean(sd_AE),2),
  sd_GM = signif(mean(sd_GM),2),
  sd_K = signif(mean(sd_K),2),
  sd_S = signif(mean(sd_S),2),
  sd_I = signif(mean(sd_I),2),
  sd_GI = signif(mean(sd_GI),2)
  # ,
  # SD_percent_T = sd_T*100/abs(meanT),
  # SD_percent_AT = sd_AT*100/abs(meanAT),
  # SD_percent_AE = sd_AE*100/abs(meanAE),
  # SD_percent_GM = sd_GM*100/abs(meanGM),
  # SD_percent_K = sd_K*100/abs(meanK),
  # SD_percent_S = sd_S*100/abs(meanS),
  # SD_percent_I = sd_I*100/abs(meanI),
  # SD_percent_GI = sd_GI*100/abs
  )
# %>%
#   kable() %>%
#   kable_styling()

# paste("Desvio padrão médio de AvgT = ", signif(mean(SD$sd_T),2), ", desvio padrao do desvio padrão = ", signif(sd(SD$sd_T),2))
# paste("Desvio padrão médio de K = ", signif(mean(SD$sd_K),2), ", desvio padrao do desvio padrão = ", signif(sd(SD$sd_K),2))
# paste("Desvio padrão médio de S = ", signif(mean(SD$sd_S),2), ", desvio padrao do desvio padrão = ", signif(sd(SD$sd_S),2))
# paste("Desvio padrão médio de I = ", signif(mean(SD$sd_I),2), ", desvio padrao do desvio padrão = ", signif(sd(SD$sd_I),2))
# paste("Desvio padrão médio de GI = ", signif(mean(SD$sd_GI),2), ", desvio padrao do desvio padrão = ", signif(sd(SD$sd_GI),2))

```

```{r}
sd_T <- ggplot(SD, aes(x = sd_T)) +
    geom_histogram(aes(alpha = 0.5)) +
  geom_vline(aes(xintercept = mean(sd_T))) +
    geom_vline(aes(xintercept = mean(sd_T) - sd(sd_T)), linetype = "dashed") +
  geom_vline(aes(xintercept = mean(sd_T) + sd(sd_T)), linetype = "dashed") +
  guides(alpha = FALSE) +
  labs(x = "SD - Average Cortical Thickness [mm]") +
    theme_pubr() + 
    theme(axis.title = element_text(size = 11),
          axis.text = element_text(size = 10), text = element_text(size = 10))

sd_GI <- ggplot(SD, aes(x = sd_GI)) +
    geom_histogram(aes(alpha = 0.5)) +
  geom_vline(aes(xintercept = mean(sd_GI))) +
    geom_vline(aes(xintercept = mean(sd_GI) - sd(sd_GI)), linetype = "dashed") +
  geom_vline(aes(xintercept = mean(sd_GI) + sd(sd_GI)), linetype = "dashed") +
  guides(alpha = FALSE) +
  labs(x = "SD - Average Cortical Thickness [mm]") +
    theme_pubr() + 
    theme(axis.title = element_text(size = 11),
          axis.text = element_text(size = 10), text = element_text(size = 10))

sd_K <- ggplot(SD, aes(x = sd_K, alpha = 0.5)) +
  geom_histogram(aes(alpha = 0.5)) +
  geom_vline(aes(xintercept = mean(sd_K))) +
      geom_vline(aes(xintercept = mean(sd_K) - sd(sd_K)), linetype = "dashed") +
  geom_vline(aes(xintercept = mean(sd_K) + sd(sd_K)), linetype = "dashed") +
  guides(alpha = FALSE) +
    labs(x = "SD - K") +
  theme_pubr() + 
  theme(axis.title = element_text(size = 11),
    axis.text = element_text(size = 10), text = element_text(size = 10))

sd_S <- ggplot(SD, aes(x = sd_S, alpha = 0.5)) +
  geom_histogram(aes(alpha = 0.5)) +
  geom_vline(aes(xintercept = mean(sd_S))) +
        geom_vline(aes(xintercept = mean(sd_S) - sd(sd_S)), linetype = "dashed") +
  geom_vline(aes(xintercept = mean(sd_S) + sd(sd_S)), linetype = "dashed") +
  guides(alpha = FALSE) +
      labs(x = "SD - S") +
  theme_pubr() + 
  theme(axis.title = element_text(size = 11),
    axis.text = element_text(size = 10), text = element_text(size = 10))

sd_I <- ggplot(SD, aes(x = sd_I, alpha = 0.5)) +
  geom_histogram(aes(alpha = 0.5)) +
  geom_vline(aes(xintercept = mean(sd_I))) +
        geom_vline(aes(xintercept = mean(sd_I) - sd(sd_I)), linetype = "dashed") +
  geom_vline(aes(xintercept = mean(sd_I) + sd(sd_I)), linetype = "dashed") +
  guides(alpha = FALSE) +
      labs(x = "SD - I") +
  theme_pubr() + 
  theme(axis.title = element_text(size = 11),
    axis.text = element_text(size = 10), text = element_text(size = 10))

sd <- ggarrange(sd_T, sd_K, sd_S, sd_I, nrow = 2, ncol = 2, labels = c("A", "B", "C", "D"), font.label = list(size = 11))

ggsave("sd.pdf", plot = sd, dpi=1200, width = 18, height = 18, units = "cm", device = "pdf")
```


```{r figure2, fig.cap="\\label{fig:figure2}Standard Deviation (SD) distribution for each variable. The solid line represents the mean value, and the dashed line the standard deviation of the distribution. (A) Average Cortical Thickness: 0.019+-0.013~mm; (B) K: 0.0017+-0.0012; (C) S: 0.014+-0.01; (D) I: 0.0064+-0.0043.", fig.height=7.1, fig.width=7.1}

sd
```

