---
title: "Typical"
author: "Fernanda Hansen P. de Moraes"
date: "05/03/2021"
output: 
  html_document: 
    toc: yes
editor_options: 
  chunk_output_type: inline
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo=FALSE,
                      message=FALSE,
                      warning=FALSE,
                      cache=TRUE)
```

```{r preparo, message=FALSE, warning=FALSE, include=FALSE}
# PREPARO

# define a area de trabalho
setwd("C:/Users/ferna/Documents/idor/Gyrification/RRRRRR/comparing_samples/") # se no computador

# carrega os pacotes utilizados
source("01-call_packages.R")

# Chama funcoes caso necessário
source("02-funcoes.R")

# onde estao os arquivos de resultados:
path <- str_c("C:/Users/ferna/Documents/idor/Gyrification/STATS/long_tables/")

path_yujiangscript <- str_c("C:/Users/ferna/Documents/idor/Gyrification/STATS/Wang_code_extraction/")

path_yujiangscript_newsurfaces <- str_c("C:/Users/ferna/Documents/idor/Gyrification/STATS/Wang_code_extraction/")
  
path_lobes <- str_c("C:/Users/ferna/Documents/idor/Gyrification/STATS/LobesScaling/")
 
path_sessions <- str_c("C:/Users/ferna/Documents/idor/Gyrification/Processamento_longitudinal/V6/data/")

Age.cor = 25

set.seed(1)

# Prepara os dados para analise
source("05-analises_prep.R")

dados_datasetscomp <- dados_datasetscomp %>%
  filter(Sample != "HCP500r", Sample != "ds000030",
         Diagnostic != "MCI")


# dados_datasetscomp <- dados_datasetscomp %>% filter(Age == 18 | Age > 18)

# write.csv(dados_datasetscomp,"dados_datasetscomp.csv")
```

# Datasets description

We included multiple datasets from different sites to estimate, accounting methodological and "subject" biases, typical values for K, S and I.

```{r description, eval=FALSE, include=FALSE}
filter(dados_datasetscomp, Sample != "Mota&Houzel2015", ROI == "hemisphere") %>%
  group_by(Diagnostic) %>%
  summarise(
    N = n_distinct(SUBJ),
    age = paste(signif(mean(Age), 2), "±", signif(sd(Age), 2)),
    age_range = paste(signif(min(Age), 2), "; ", signif(max(Age), 2))
) %>% kable(digits = 2) %>% kable_styling()

filter(dados_datasetscomp, Sample != "Mota&Houzel2015", ROI == "hemisphere") %>%
  group_by(Diagnostic, Gender) %>%
  summarise(
    N = n_distinct(SUBJ)
) %>% kable(digits = 2) %>% kable_styling()

filter(dados_datasetscomp, Sample != "Mota&Houzel2015", ROI == "hemisphere") %>%
  summarise(
    N = n_distinct(SUBJ),
    age = paste(signif(mean(Age), 2), "±", signif(sd(Age), 2)),
    age_range = paste(signif(min(Age), 2), "; ", signif(max(Age), 2))
) %>% kable(digits = 2) %>% kable_styling()

filter(dados_datasetscomp,
       Sample != "Mota&Houzel2015",
       ROI == "hemisphere") %>%
  group_by(Sample, Diagnostic) %>%
  summarise(
    N = n_distinct(SUBJ),
    age = paste(signif(mean(Age), 2), "±", signif(sd(Age), 2)),
    age_range = paste(signif(min(Age), 2), "; ", signif(max(Age), 2)),
    AvgT =  paste(signif(mean(AvgThickness), 2), "±", signif(sd(AvgThickness), 2)),
    AT =  paste(signif(mean(TotalArea), 2), "±", signif(sd(TotalArea), 2)),
    AE =  paste(signif(mean(ExposedArea), 2), "±", signif(sd(ExposedArea), 2)),
    K =  paste(signif(mean(K), 2), "±", signif(sd(K), 2)),
    S =  paste(signif(mean(S), 2), "±", signif(sd(S), 2)),
    I =  paste(signif(mean(I), 2), "±", signif(sd(I), 2))
  ) %>% kable(digits = 2) %>% kable_styling()


filter(dados_datasetscomp, Sample != "Mota&Houzel2015", ROI == "hemisphere") %>%
  group_by(Sample, Diagnostic, Gender) %>%
  summarise(
    N = n_distinct(SUBJ)
) %>% kable(digits = 2) %>% kable_styling()
```
# Mapping

# correction factor lobes

```{r eval=FALSE, include=FALSE}
ggplot(filter(dados_datasetscomp, ROI != "hemisphere"), aes(x = ROI, y = c, color = Sample)) +
  geom_boxplot() +
  geom_jitter() +
  theme_pubr()
```

## CTL

```{r}
ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = logExposedArea, y = 1/2 * logAvgThickness + logTotalArea)) +
  geom_point(aes(color = Sample)) +
  geom_smooth()+
  theme_pubr()

amostra_Coef <- filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL", Age_interval != "[0,5)", Age_interval != "[95,100]") %>%
    group_by(Age_interval) %>%
    do(fit_amostra = tidy(lm(1/2 * logAvgThickness + logTotalArea ~ logExposedArea, data = ., na.action = na.omit), conf.int = TRUE, conf.level = 0.95)) %>% 
    unnest(fit_amostra)

amostras_Coef_age <- filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL", Age_interval != "[0,5)", Age_interval != "[95,100]") %>%
    group_by(Age_interval) %>%
    summarise(
        N = n_distinct(SUBJ),
        min_age = min(Age),
        max_age = max(Age))

amostra_Coef <- full_join(amostra_Coef, amostras_Coef_age) %>% filter(term == "logExposedArea")

amostra_Coef$Age_intervalnumber <- as.numeric(factor(amostra_Coef$Age_interval))-1

cor.test(amostra_Coef$estimate, amostra_Coef$Age_intervalnumber)

cor.test(filter(amostra_Coef, Age_intervalnumber > 1)$estimate, filter(amostra_Coef, Age_intervalnumber > 1)$Age_intervalnumber)

ggplot(amostra_Coef, aes(y = estimate, x = Age_interval)) +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
  geom_point() +
  geom_hline(yintercept = 1.25, linetype = "dashed") +
  theme_pubr() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=9)
  ) +
  labs(y ="Slope", x = "Age interval") +
  geom_text(aes(label = N), nudge_y = 0.3) +
  geom_line(group = 0 )

amostra_Coef %>% dplyr::select(-c(term)) %>% kable() %>% kable_styling()

ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL", Age_interval != "[0,5)", Age_interval != "[95,100]"), aes(y = K, x = Age_interval), aes(y = K, x = Age_interval)) +
    geom_boxplot() + 
    theme_pubr() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=9)
    ) 
```

```{r}
ggplot(filter(dados_datasetscomp, ROI != "hemisphere", Diagnostic == "CTL"), aes(x = logExposedArea_corrected, y = 1/2 * logAvgThickness + logTotalArea_corrected)) +
    geom_point(aes(color = ROI)) +
    geom_smooth(method = "lm")+
    theme_pubr() + facet_grid(Sample ~ .)

amostra_Coef <- filter(dados_datasetscomp, ROI != "hemisphere", Diagnostic == "CTL", Age_interval != "[0,5)", Age_interval != "[95,100]") %>%
    group_by(Age_interval, ROI) %>%
    do(fit_amostra = tidy(lm(1/2 * logAvgThickness + logTotalArea_corrected ~ logExposedArea_corrected, data = ., na.action = na.omit), conf.int = TRUE, conf.level = 0.95)) %>% unnest(fit_amostra)

amostras_Coef_age <- filter(dados_datasetscomp, ROI != "hemisphere", Diagnostic == "CTL", Age_interval != "[0,5)", Age_interval != "[95,100]") %>%
    group_by(Age_interval) %>%
    summarise(
        N = n_distinct(SUBJ),
        min_age = min(Age),
        max_age = max(Age))

amostra_Coef <- full_join(amostra_Coef, amostras_Coef_age) %>% filter(term == "logExposedArea_corrected")

ggplot(amostra_Coef, aes(y = estimate, x = Age_interval)) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
    geom_point() +
  geom_hline(yintercept = 1.25, linetype = "dashed") +
    theme_pubr() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=9)) +
  labs(y ="Slope", x = "Age interval") +
  geom_text(aes(label = N), nudge_y = 0.3) +
  facet_grid(ROI ~ .)

amostra_Coef %>%
  dplyr::select(-c(term)) %>%
  kable() %>%
  kable_styling()

ggplot(filter(dados_datasetscomp, ROI != "hemisphere", Diagnostic == "CTL", Age_interval != "[0,5)", Age_interval != "[95,100]"), aes(y = K, x = Age_interval), aes(y = K, x = Age_interval)) +
    geom_boxplot() + 
    theme_pubr() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=9)) +
  facet_grid(ROI ~ .)

```

### per subject
```{r eval=FALSE, include=FALSE}
amostra_Coef <- filter(dados_datasetscomp, ROI != "hemisphere", Diagnostic == "CTL", Age_interval != "[0,5)", Age_interval != "[95,100]") %>%
  group_by(SUBJ) %>%
  do(fit_amostra = tidy(lm(1/2 * logAvgThickness + logTotalArea_corrected ~ logExposedArea_corrected, data = ., na.action = na.omit), conf.int = TRUE, conf.level = 0.95)) %>% 
  unnest(fit_amostra)

ggplot(filter(amostra_Coef, term == "logExposedArea_corrected"), aes(x = estimate, alpha = 0.3)) +
  geom_histogram() +
  geom_vline(xintercept = 1.25, linetype = "dashed") +
  theme_pubr() + 
  theme(axis.text.x = element_text(angle = 45, hjust = 1, size=9)) +
  labs(x ="Slope", x = "Age interval") +
  guides(alpha = "none")
```

```{r}
amostra_Coef <- filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL" | Diagnostic == "AD", Age_interval != "[0,5)", Age_interval != "[95,100]") %>%
    group_by(Age_interval, Diagnostic) %>%
    do(fit_amostra = tidy(lm(1/2 * logAvgThickness + logTotalArea ~ logExposedArea, data = ., na.action = na.omit), conf.int = TRUE, conf.level = 0.95)) %>% 
    unnest(fit_amostra)

amostras_Coef_age <- filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL" | Diagnostic == "AD", Age_interval != "[0,5)", Age_interval != "[95,100]") %>%
    group_by(Age_interval, Diagnostic) %>%
    summarise(
        N = n_distinct(SUBJ),
        min_age = min(Age),
        max_age = max(Age))

amostra_Coef <- full_join(amostra_Coef, amostras_Coef_age) %>% filter(term == "logExposedArea")

amostra_Coef$Diagnostic <- factor(amostra_Coef$Diagnostic, levels=c('AD', 'CTL'))

ggplot(amostra_Coef, aes(y = estimate, x = Age_interval, color = Diagnostic)) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
    geom_point() + geom_hline(yintercept = 1.25, linetype = "dashed") +
    theme_pubr() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=9)
    ) + labs(y ="Slope", x = "Age interval") +  geom_text(aes(label = N), nudge_y = 0.3)

amostra_Coef %>% dplyr::select(-c(term)) %>% kable() %>% kable_styling()

```

```{r mapping}
paste("Total number of CTL subjects =", n_distinct(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL")$SUBJ))
paste("Age range CTL. Min age = ", min(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL")$Age), "; max age = ", max(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL")$Age))

figS6a <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"),
       aes(x = Age , y = K)) +
  geom_point(aes(color = Sample, fill = Sample, alpha = 0.4)) +
  geom_smooth(color = "black", method = "lm") +
  theme_pubr() +
  guides(alpha = "none")+ 
  labs(x = "Age [years]")

# ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"),
#        aes(x = (Age*12*4)+40 , y = K)) +
#     geom_point(aes(color = Sample, fill = Sample, alpha = 0.4)) +
#     geom_smooth(color = "black", method = "lm") + geom_vline(xintercept = 40, linetype = "dashed") +
#     theme_pubr() +
#     guides(alpha = "none")+ 
#     labs(x = "Age [weeks] (age of birth defined as 40 weeks)") + xlim(0,1000)

#figS6a

figS6b <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"),
       aes(x = Age, y = S)) +
  geom_point(aes(color = Sample, fill = Sample, alpha = 0.4)) +
  geom_smooth(color = "black", method = "lm") +
  theme_pubr() +
  guides(alpha = "none", color = FALSE,fill = FALSE)+ 
  labs(x = "Age [years]") +
  theme(legend.position= "none")

#figS6b

figS6c <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"),
       aes(x = Age, y = I)) + geom_point(aes(color = Sample, fill = Sample, alpha = 0.4)) + 
  geom_smooth(color = "black", method = "lm") + theme_pubr() +
  guides(alpha = "none", color = FALSE,fill = FALSE)+ 
  labs(x = "Age [years]") +
  theme(legend.position= "none")
 
# figS6c      

figS6 <- ggarrange(figS6a, ggarrange(figS6b, figS6c,labels = c("B","C"), nrow = 1, ncol = 2, font.label = list(size = 11)),labels = c("A"), nrow = 2, ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")

# figS6

ggsave("figS6.pdf", plot = figS6, dpi=1200, width = 18, height = 22, units = "cm", device = "pdf")
# ggsave("figS6.png", plot = figS6, dpi=1200, width = 18, height = 22, units = "cm", device = "png")

cor.test(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL")$K, filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL")$Age)

cor.test(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL")$S, filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL")$Age)

cor.test(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL")$I, filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL")$Age)
```

# Decreasing rate

```{r}
dados_datasetscomp_rate <- filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL" | Diagnostic == "MCI"| Diagnostic == "AD", Sample != "IDOR-CCD-Control")
dados_datasetscomp_rate$Diagnostic <- factor(dados_datasetscomp_rate$Diagnostic, levels = c("CTL", "MCI","AD"))
dados_datasetscomp_rate$Sample <- as.factor(dados_datasetscomp_rate$Sample)

N_SUBJ <- dados_datasetscomp_rate %>% group_by(Diagnostic) %>% summarise(N_SUBJ = n_distinct(SUBJ))
```

## AvgThickness \~ Age

```{r cache=TRUE}
m.1 <- lme4::lmer(AvgThickness ~ Age * Diagnostic +(1|Sample) + (1|Sample:Diagnostic) , data = filter(dados_datasetscomp, ROI == "hemisphere"))

## performance::check_model(m.1)

anova(m.1)
sigma.hat(m.1)
r.squaredGLMM(m.1)

mean <- lsmeans(m.1, ~ Diagnostic, var="Age")
m.lstk <- lstrends(m.1, ~ Diagnostic, var="Age")
pairs <- pairs(m.lstk) 

mean <- as_tibble(mean)
m.lstk <- as_tibble(m.lstk)

tableT <- full_join(mean, m.lstk, by = c("Diagnostic")) %>% mutate(percent = Age.trend*100/abs(lsmean), Variable = "T")

# m.lstk %>% mutate(
#     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)),
#     CI = paste(signif(asymp.LCL, 2), ";", signif(asymp.UCL, 2)),
#   ) %>% dplyr::select(c(Diagnostic, TX, CI)) %>% full_join(N_SUBJ)
# 
# m.lstk %>% kable() %>% kable_styling()

coefs <- data.frame(coef(summary(m.1)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
# coefs

# pairs %>% kable() %>% kable_styling()
 
# ggplot(data = as_tibble(pairs), aes(
#     x = contrast,
#     y = estimate,
#     ymin = estimate - SE,
#     ymax = estimate + SE)) +
#     geom_hline(yintercept = 0,
#                linetype = "11",
#                colour = "grey60") +
#     geom_pointrange( position = position_dodge(width = 0.2)) +
#    geom_text(aes(label = str_c("p adj ", signif(p.value, digits = 2))), nudge_x = 0.3, nudge_y = 0.0003) +
#     coord_flip() + 
#     labs(y =  expression('Differences in slopes ('*Delta*'AvgT/'*Delta*'Age)'), x = "Contrast") +
#   theme_pubr()

# e <- effect("Age:Diagnostic", m.1)
# e <- as.data.frame(e)
# 
# ggplot(e, aes(x = Age, y = fit,ymin=lower, ymax=upper, shape = Diagnostic)) + geom_pointrange() + geom_line() + theme_pubr() + labs(x = "Age [years]", y = "Avg Cortical Thickness", shape = "Diagnostic")

# re <- as_tibble(ranef(m.1))

# ggplot(filter(re, grpvar == "Sample"), aes(y = condval, x = as.character(grp))) +
#   geom_errorbar(aes(ymin = condval+1.96*condsd, ymax = condval-1.96*condsd), width = 0.2) +
#   geom_point() + geom_hline(yintercept = 0, linetype = "dashed") +
#   theme_pubr() + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10),
#     text = element_text(size = 10),
#     axis.text.x = element_text(angle = 45, hjust = 1, size=9)
#   ) + labs(y ="T variation", x = "Sample")

```

## TotalArea \~ Age

```{r eval=FALSE, cache=TRUE, include=FALSE}
m.1 <- lme4::lmer(TotalArea ~ Age * Diagnostic +(1|Sample) + (1|Sample:Diagnostic) , data = filter(dados_datasetscomp, ROI == "hemisphere"))

# performance::check_model(m.1)

anova(m.1)
summary(m.1)
s <- summary(m.1)
sigma.hat(m.1)
r.squaredGLMM(m.1)

mean <- lsmeans(m.1, ~ Diagnostic, var="Age")
m.lstk <- lstrends(m.1, ~ Diagnostic, var="Age")
pairs <- pairs(m.lstk) 

mean <- as_tibble(mean)
m.lstk <- as_tibble(m.lstk)

tableAT <- full_join(mean, m.lstk, by = c("Diagnostic")) %>% mutate(percent = Age.trend*100/abs(lsmean), Variable = "AT")

m.lstk %>% mutate(
    TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)),
    CI = paste(signif(asymp.LCL, 2), ";", signif(asymp.UCL, 2)),
  ) %>% dplyr::select(c(Diagnostic, TX, CI)) %>% full_join(N_SUBJ)

m.lstk %>% kable() %>% kable_styling()

coefs <- data.frame(coef(summary(m.1)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
# coefs

# pairs %>% kable() %>% kable_styling()
 
# ggplot(data = as_tibble(pairs), aes(
#     x = contrast,
#     y = estimate,
#     ymin = estimate - SE,
#     ymax = estimate + SE)) +
#     geom_hline(yintercept = 0,
#                linetype = "11",
#                colour = "grey60") +
#     geom_pointrange( position = position_dodge(width = 0.2)) +
#    geom_text(aes(label = str_c("p adj ", signif(p.value, digits = 2))), nudge_x = 0.3, nudge_y = 0.0003) +
#     coord_flip() + 
#     labs(y =  expression('Differences in slopes ('*Delta*'AT/'*Delta*'Age)'), x = "Contrast") +
#   theme_pubr() 

# e <- effect("Age:Diagnostic", m.1)
# e <- as.data.frame(e)
 
# ggplot(e, aes(x = Age, y = fit,ymin=lower, ymax=upper, shape = Diagnostic)) + geom_pointrange() + geom_line() + theme_pubr() + labs(x = "Age [years]", y = "AT", shape = "Diagnostic") 

```

## ExposedArea \~ Age

```{r eval=FALSE, cache=TRUE, include=FALSE}
m.1 <- lme4::lmer(ExposedArea ~ Age * Diagnostic +(1|Sample) + (1|Sample:Diagnostic) , data = filter(dados_datasetscomp, ROI == "hemisphere"))

# performance::check_model(m.1)

anova(m.1)
summary(m.1)
s <- summary(m.1)
sigma.hat(m.1)
r.squaredGLMM(m.1)

mean <- lsmeans(m.1, ~ Diagnostic, var="Age")
m.lstk <- lstrends(m.1, ~ Diagnostic, var="Age")
pairs <- pairs(m.lstk) 

mean <- as_tibble(mean)
m.lstk <- as_tibble(m.lstk)

tableAE <- full_join(mean, m.lstk, by = c("Diagnostic")) %>% mutate(percent = Age.trend*100/abs(lsmean), Variable = "AE")

m.lstk %>% mutate(
    TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)),
    CI = paste(signif(asymp.LCL, 2), ";", signif(asymp.UCL, 2)),
  ) %>% dplyr::select(c(Diagnostic, TX, CI)) %>% full_join(N_SUBJ)

m.lstk %>% kable() %>% kable_styling()

# coefs <- data.frame(coef(summary(m.1)))
# coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
# coefs

# pairs %>% kable() %>% kable_styling()
 
# ggplot(data = as_tibble(pairs), aes(
#     x = contrast,
#     y = estimate,
#     ymin = estimate - SE,
#     ymax = estimate + SE)) +
#     geom_hline(yintercept = 0,
#                linetype = "11",
#                colour = "grey60") +
#     geom_pointrange( position = position_dodge(width = 0.2)) +
#    geom_text(aes(label = str_c("p adj ", signif(p.value, digits = 2))), nudge_x = 0.3, nudge_y = 0.0003) +
#     coord_flip() + 
#     labs(y =  expression('Differences in slopes ('*Delta*'AE/'*Delta*'Age)'), x = "Contrast") +
#   theme_pubr()

# e <- effect("Age:Diagnostic", m.1)
# e <- as.data.frame(e)

# ggplot(e, aes(x = Age, y = fit,ymin=lower, ymax=upper, shape = Diagnostic)) + geom_pointrange() + geom_line() + theme_pubr() + labs(x = "Age [years]", y = "AE", shape = "Diagnostic") 

```

## GI \~ Age

```{r cache=TRUE}
m.1 <- lme4::lmer(localGI ~ Age * Diagnostic + (1|Sample) + (1|Sample:Diagnostic), data = filter(dados_datasetscomp, ROI == "hemisphere"))

# performance::check_model(m.1)

anova(m.1)
sigma.hat(m.1)
r.squaredGLMM(m.1)

mean <- lsmeans(m.1, ~ Diagnostic, var="Age")
m.lstk <- lstrends(m.1, ~ Diagnostic, var="Age")
pairs <- pairs(m.lstk) 

mean <- as_tibble(mean)
m.lstk <- as_tibble(m.lstk)

tableGI <- full_join(mean, m.lstk, by = c("Diagnostic")) %>% mutate(percent = Age.trend*100/abs(lsmean), Variable = "GI")

# m.lstk %>% mutate(
#     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)),
#     CI = paste(signif(asymp.LCL, 2), ";", signif(asymp.UCL, 2)),
#   ) %>% dplyr::select(c(Diagnostic, TX, CI)) %>% full_join(N_SUBJ)
# 
# m.lstk %>% kable() %>% kable_styling()

coefs <- data.frame(coef(summary(m.1)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
# coefs

# pairs %>% kable() %>% kable_styling()
 
# ggplot(data = as_tibble(pairs), aes(
#     x = contrast,
#     y = estimate,
#     ymin = estimate - SE,
#     ymax = estimate + SE)) +
#     geom_hline(yintercept = 0,
#                linetype = "11",
#                colour = "grey60") +
#     geom_pointrange( position = position_dodge(width = 0.2)) +
#    geom_text(aes(label = str_c("p adj ", signif(p.value, digits = 2))), nudge_x = 0.3, nudge_y = 0.0003) +
#     coord_flip() + 
#     labs(y =  expression('Differences in slopes ('*Delta*'GI/'*Delta*'Age)'), x = "Contrast") +
#   theme_pubr() 

# e <- effect("Age:Diagnostic", m.1)
# e <- as.data.frame(e)

# ggplot(e, aes(x = Age, y = fit,ymin=lower, ymax=upper, shape = Diagnostic)) + geom_pointrange() + geom_line() + theme_pubr() + labs(x = "Age [years]", y = "GI", shape = "Diagnostic") 

```


## K \~ Age

```{r cache=TRUE}
N_SUBJ <- dados_datasetscomp_rate %>% group_by(Diagnostic) %>% summarise(N_SUBJ = n_distinct(SUBJ))

m.1 <- lme4::lmer(K ~ Age * Diagnostic + (1|Sample) + (1|Sample:Diagnostic), data = filter(dados_datasetscomp, ROI == "hemisphere"))

# performance::check_model(m.1)

anova(m.1)
sigma.hat(m.1)
r.squaredGLMM(m.1)

mean <- lsmeans(m.1, ~ Diagnostic, var="Age")
m.lstk <- lstrends(m.1, ~ Diagnostic, var="Age")
pairs <- pairs(m.lstk) 

mean <- as_tibble(mean)
m.lstk <- as_tibble(m.lstk)

tableK <- full_join(mean, m.lstk, by = c("Diagnostic")) %>% mutate(percent = Age.trend*100/abs(lsmean), Variable = "K")

# m.lstk %>% mutate(
#     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)),
#     CI = paste(signif(asymp.LCL, 2), ";", signif(asymp.UCL, 2)),
#   ) %>% dplyr::select(c(Diagnostic, TX, CI)) %>% full_join(N_SUBJ) 
# m.lstk %>% kable() %>% kable_styling()

coefs <- data.frame(coef(summary(m.1)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
# coefs

# pairs %>% kable() %>% kable_styling()
 
fig_TX_K_age <- ggplot(data = as_tibble(pairs), aes(
    x = contrast,
    y = estimate,
    ymin = estimate - SE,
    ymax = estimate + SE)) +
    geom_hline(yintercept = 0,
               linetype = "11",
               colour = "grey60") +
    geom_pointrange( position = position_dodge(width = 0.2)) +
   geom_text(aes(label = str_c("p adj ", signif(p.value, digits = 2))), nudge_x = 0.3, nudge_y = 0.0003) +
    coord_flip() + 
    labs(y =  expression('Differences in slopes ('*Delta*'K/'*Delta*'Age)'), x = "Contrast") +
  theme_pubr()

e <- effect("Age:Diagnostic", m.1)
e <- as.data.frame(e)
fig_K_age <- ggplot(e, aes(x = Age, y = fit,ymin=lower, ymax=upper, shape = Diagnostic)) + geom_pointrange() + geom_line() + theme_pubr() + labs(x = "Age [years]", y = "K", shape = "Diagnostic")

# re <- as_tibble(ranef(m.1))

# ggplot(filter(re, grpvar == "Sample"), aes(y = condval, x = as.character(grp))) +
#   geom_errorbar(aes(ymin = condval+1.96*condsd, ymax = condval-1.96*condsd), width = 0.2) +
#   geom_point() + geom_hline(yintercept = 0, linetype = "dashed") +
#   theme_pubr() + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10),
#     text = element_text(size = 10),
#     axis.text.x = element_text(angle = 45, hjust = 1, size=9)
#   ) + labs(y ="K variation", x = "Sample")

```


## S \~ Age

```{r cache=TRUE}
m.1 <- lme4::lmer(S ~ Age * Diagnostic +(1|Sample) + (1|Sample:Diagnostic) , data = filter(dados_datasetscomp, ROI == "hemisphere"))

# performance::check_model(m.1)

anova(m.1)
sigma.hat(m.1)
r.squaredGLMM(m.1)

mean <- lsmeans(m.1, ~ Diagnostic, var="Age")
m.lstk <- lstrends(m.1, ~ Diagnostic, var="Age")
pairs <- pairs(m.lstk) 

mean <- as_tibble(mean)
m.lstk <- as_tibble(m.lstk)

tableS <- full_join(mean, m.lstk, by = c("Diagnostic")) %>% mutate(percent = Age.trend*100/abs(lsmean), Variable = "S")

# m.lstk %>% mutate(
#     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)),
#     CI = paste(signif(asymp.LCL, 2), ";", signif(asymp.UCL, 2)),
#   ) %>% dplyr::select(c(Diagnostic, TX, CI)) %>% full_join(N_SUBJ)
# m.lstk %>% kable() %>% kable_styling()

coefs <- data.frame(coef(summary(m.1)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
# coefs

# pairs %>% kable() %>% kable_styling()
 
fig_TX_S_age <- ggplot(data = as_tibble(pairs), aes(
    x = contrast,
    y = estimate,
    ymin = estimate - SE,
    ymax = estimate + SE)) +
    geom_hline(yintercept = 0,
               linetype = "11",
               colour = "grey60") +
    geom_pointrange( position = position_dodge(width = 0.2)) +
   geom_text(aes(label = str_c("p adj ", signif(p.value, digits = 2))), nudge_x = 0.3, nudge_y = 0.0003) +
    coord_flip() + 
    labs(y =  expression('Differences in slopes ('*Delta*'S/'*Delta*'Age)'), x = "Contrast") +
  theme_pubr() + 
  theme(axis.title = element_text(size = 11),
    axis.text = element_text(size = 10), text = element_text(size = 10))

e <- effect("Age:Diagnostic", m.1)
e <- as.data.frame(e)
fig_S_age <- ggplot(e, aes(x = Age, y = fit,ymin=lower, ymax=upper, shape = Diagnostic)) + geom_pointrange() + geom_line() + theme_pubr() + labs(x = "Age [years]", y = "S", shape = "Diagnostic") + 
  theme(legend.position = "none")

```

## I \~ Age

```{r cache=TRUE}
m.1 <- lme4::lmer(I ~ Age * Diagnostic +(1|Sample) + (1|Sample:Diagnostic) , data = filter(dados_datasetscomp, ROI == "hemisphere"))

# performance::check_model(m.1)

anova(m.1)
sigma.hat(m.1)
r.squaredGLMM(m.1)

mean <- lsmeans(m.1, ~ Diagnostic, var="Age")
m.lstk <- lstrends(m.1, ~ Diagnostic, var="Age")
pairs <- pairs(m.lstk) 

mean <- as_tibble(mean)
m.lstk <- as_tibble(m.lstk)

tableI <- full_join(mean, m.lstk, by = c("Diagnostic")) %>% mutate(percent = Age.trend*100/abs(lsmean), Variable = "I")

# m.lstk %>% mutate(
#     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)),
#     CI = paste(signif(asymp.LCL, 2), ";", signif(asymp.UCL, 2)),
#   ) %>% dplyr::select(c(Diagnostic, TX, CI)) %>% full_join(N_SUBJ)
# m.lstk %>% kable() %>% kable_styling()

coefs <- data.frame(coef(summary(m.1)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
# coefs

# pairs %>% kable() %>% kable_styling()
 
fig_TX_I_age <- ggplot(data = as_tibble(pairs), aes(
    x = contrast,
    y = estimate,
    ymin = estimate - SE,
    ymax = estimate + SE)) +
    geom_hline(yintercept = 0,
               linetype = "11",
               colour = "grey60") +
    geom_pointrange( position = position_dodge(width = 0.2)) +
   geom_text(aes(label = str_c("p adj ", signif(p.value, digits = 2))), nudge_x = 0.3, nudge_y = 0.0003) +
    coord_flip() + 
    labs(y =  expression('Differences in slopes ('*Delta*'I/'*Delta*'Age)'), x = "Contrast") +
  theme_pubr() + 
  theme(text = element_text(size = 10))

e <- effect("Age:Diagnostic", m.1)
e <- as.data.frame(e)

fig_I_age <- ggplot(e, aes(x = Age, y = fit,ymin=lower, ymax=upper, shape = Diagnostic)) + geom_pointrange() + geom_line() + theme_pubr() + labs(x = "Age [years]", y = "I", shape = "Diagnostic") + 
  theme(legend.position = "none")

```

## rates

```{r echo=FALSE}
tableT <- tableT %>% dplyr::select(c(Diagnostic, percent, Variable, Age.trend, SE.y))
tableK <- tableK %>% dplyr::select(c(Diagnostic, percent, Variable, Age.trend, SE.y))
tableS <- tableS %>% dplyr::select(c(Diagnostic, percent, Variable, Age.trend, SE.y))
tableI <- tableI %>% dplyr::select(c(Diagnostic, percent, Variable, Age.trend, SE.y))
tableGI <- tableGI %>% dplyr::select(c(Diagnostic, percent, Variable, Age.trend, SE.y))

rate <- full_join(tableT, tableK) %>%
  full_join(tableS) %>%
  full_join(tableI) %>%
  full_join(tableGI) %>%
  mutate(TX = paste(signif(Age.trend, 2), "±", signif(SE.y, 2)), percent = signif(percent, 2))
  
rate %>%
  kable() %>%
  kable_styling()

```

# Harmonization

# Figures

```{r eval=FALSE, include=FALSE}
T_shifta <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = AvgThickness, color = Sample))+
  geom_point() +
  theme_pubr() +
  labs(y = "T") +
  scale_x_continuous(limits = c(0,100))

T_shiftb <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = AvgThickness_shiftc, color = Sample)) +
  geom_point() +
  theme_pubr() +
  labs(y = "T (harmonized)") +
  scale_x_continuous(limits = c(0,100))

T_shift <- ggarrange(T_shifta, T_shiftb,labels = c("A","B"), nrow = 2, ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")
# T_shift
# ggsave("GI_shift.pdf", plot = GI_shift, width = 18, height = 22, units = "cm", device = "pdf")
```

```{r eval=FALSE, include=FALSE}
GI_shifta <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = localGI, color = Sample))+
  geom_point() +
  theme_pubr() +
  labs(y = "GI") +
  scale_x_continuous(limits = c(0,100))

GI_shiftb <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = localGI_shiftc, color = Sample)) +
  geom_point() +
  theme_pubr() +
  labs(y = "GI (harmonized)") +
  scale_x_continuous(limits = c(0,100))

GI_shift <- ggarrange(GI_shifta, GI_shiftb,labels = c("A","B"), nrow = 2, ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")
# GI_shift
# ggsave("GI_shift.pdf", plot = GI_shift, width = 18, height = 22, units = "cm", device = "pdf")
```

```{r echo=FALSE}
K_shifta <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = K, color = Sample, alpha = 0.4))+
  geom_point() +
  theme_pubr() + 
  guides(alpha = "none") +
  scale_x_continuous(limits = c(0,100))

K_lifespan_a <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = K, alpha = 0.4))+
    geom_point(aes(color = Sample, fill = Sample, alpha = 0.1)) +
    geom_smooth(method = "lm", color = "black") + geom_smooth(aes(color = Sample, fill = Sample), method = "lm") +
    theme_pubr() + 
    guides(alpha = "none") +
  scale_x_continuous(limits = c(0,100))

K_shiftb <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = K_shiftc, color = Sample, alpha = 0.4))+
  geom_point() +
    guides(alpha = "none") + 
  theme_pubr() +
  labs(y = "K (harmonized)") +
  scale_x_continuous(limits = c(0,100))

K_lifespan_b <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = K_shiftc, alpha = 0.4))+
    geom_point(aes(color = Sample, fill = Sample, alpha = 0.1)) +
    geom_smooth(method = "lm", color = "black") + geom_smooth(aes(color = Sample, fill = Sample), method = "lm") +
    guides(alpha = "none") + 
  theme_pubr() +
  labs(y = "K (harmonized)") +
  scale_x_continuous(limits = c(0,100))

K_shift <- ggarrange(K_shifta, K_shiftb,labels = c("A","B"), nrow = 2, ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")
#K_shift
ggsave("K_shift.pdf", plot = K_shift, width = 18, height = 22, units = "cm", device = "pdf")

K_lifespan <- ggarrange(K_lifespan_a, K_lifespan_b,labels = c("A","B"), nrow = 2, ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")
# K_lifespan
ggsave("K_lifespan.pdf", plot = K_lifespan, width = 18, height = 22, units = "cm", device = "pdf")
# ggsave("K_lifespan.png", plot = K_lifespan, width = 18, height = 22, units = "cm", device = "png")
```

```{r include=FALSE}
S_shifta <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = S, color = Sample))+
  geom_point() +
  theme_pubr() +
  scale_x_continuous(limits = c(0,100))

S_shiftb <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = S_shiftc, color = Sample))+
  geom_point() +
  theme_pubr() +
  labs(y = "S (harmonized)") +
  scale_x_continuous(limits = c(0,100))

S_lifespan_a <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = S, alpha = 0.4)) +
  geom_point(aes(color = Sample, fill = Sample, alpha = 0.1)) +
  geom_smooth(method = "lm", color = "black") +
  geom_smooth(aes(color = Sample, fill = Sample), method = "lm") +
  guides(alpha = "none") + 
  theme_pubr() +
  labs(y = "S") +
  scale_x_continuous(limits = c(0,100))

S_shift <- ggarrange(S_shifta, S_shiftb,labels = c("A","B"), nrow = 2, ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")
# S_shift
# ggsave("S_shift.pdf", plot = S_shift, width = 18, height = 22, units = "cm", device = "pdf")

S_lifespan_b <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = S_shiftc, alpha = 0.4))+
  geom_point(aes(color = Sample, fill = Sample, alpha = 0.1)) +
  geom_smooth(method = "lm", color = "black") +
  geom_smooth(aes(color = Sample, fill = Sample), method = "lm") +
  guides(alpha = "none") + 
  theme_pubr() +
  labs(y = "S (harmonized)") +
  scale_x_continuous(limits = c(0,100))

S_lifespan <- ggarrange(S_lifespan_a, S_lifespan_b,labels = c("A","B"), nrow = 2, ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")
# S_lifespan
ggsave("S_lifespan.pdf", plot = S_lifespan, width = 18, height = 22, units = "cm", device = "pdf")
```

```{r include=FALSE}
I_shifta <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = I, color = Sample))+
  geom_point() +
  theme_pubr() +
  scale_x_continuous(limits = c(0,100))

I_lifespan_a <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = I, alpha = 0.4))+
  geom_point(aes(color = Sample, fill = Sample, alpha = 0.1)) +
  geom_smooth(method = "lm", color = "black") +
  geom_smooth(aes(color = Sample, fill = Sample), method = "lm") +
  guides(alpha = "none") + 
  theme_pubr() +
  labs(y = "I") +
  scale_x_continuous(limits = c(0,100))

I_lifespan_b <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = I_shiftc, alpha = 0.4))+
  geom_point(aes(color = Sample, fill = Sample, alpha = 0.1)) +
  geom_smooth(method = "lm", color = "black") +
  geom_smooth(aes(color = Sample, fill = Sample), method = "lm") +
  guides(alpha = "none") + 
  theme_pubr() +
  labs(y = "I (harmonized)") +
  scale_x_continuous(limits = c(0,100))

I_shiftb <- ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"), aes(x = Age, y = I_shiftc, color = Sample))+
  geom_point() +
  theme_pubr() +
  labs(y = "I (harmonized)") +
  scale_x_continuous(limits = c(0,100))


I_shift <- ggarrange(I_shifta, I_shiftb,labels = c("A","B"), nrow = 2, ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")
# I_shift
# ggsave("I_shift.pdf", plot = I_shift, width = 18, height = 22, units = "cm", device = "pdf")


I_lifespan <- ggarrange(I_lifespan_a, I_lifespan_b,labels = c("A","B"), nrow = 2, ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")
# I_lifespan
ggsave("I_lifespan.pdf", plot = I_lifespan, width = 18, height = 22, units = "cm", device = "pdf")

```


```{r}
fig_TX_age <- ggarrange(fig_TX_K_age, ggarrange(fig_TX_S_age, fig_TX_I_age,labels = c("B","C"), nrow = 1, ncol = 2, font.label = list(size = 11)),labels = c("A"), nrow = 2, ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")

# fig_TX_age

fig_age <- ggarrange(fig_K_age, ggarrange(fig_S_age, fig_I_age,labels = c("B","C"), nrow = 1, ncol = 2, font.label = list(size = 11)),labels = c("A"), nrow = 2, ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")

# fig_age

# ggsave("fig_TX_age.pdf", plot = fig_TX_age, dpi=1200, width = 18, height = 18, units = "cm", device = "pdf")
# ggsave("fig_TX_age.png", plot = fig_TX_age, dpi=1200, width = 18, height = 18, units = "cm", device = "png")

# ggsave("fig_age.pdf", plot = fig_age, dpi=1200, width = 18, height = 18, units = "cm", device = "pdf")
# ggsave("fig_age.png", plot = fig_age, dpi=1200, width = 18, height = 18, units = "cm", device = "png")

# ggsave("fig_sampleeff.png", plot = fig_sampleeff, dpi=1200, width = 18, height = 18, units = "cm", device = "png")
```

# After harmonization

```{r cache=TRUE}
amostra_Coef <- filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL", Age_interval != "[0,5)", Age_interval != "[95,100]") %>%
    group_by(Age_interval) %>%
    do(fit_amostra = tidy(lm(1/2 * log10(AvgThickness_shiftc) + log10(TotalArea_shiftc) ~ log10(ExposedArea_shiftc), data = ., na.action = na.omit), conf.int = TRUE, conf.level = 0.95)) %>% 
    unnest(fit_amostra)

amostras_Coef_age <- filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL", Age_interval != "[0,5)", Age_interval != "[95,100]") %>%
    group_by(Age_interval) %>%
    summarise(
        N = n_distinct(SUBJ),
        min_age = min(Age),
        max_age = max(Age))

amostra_Coef <- full_join(amostra_Coef, amostras_Coef_age) %>% filter(term == "log10(ExposedArea_shiftc)")

amostra_Coef$Age_intervalnumber <- as.numeric(factor(amostra_Coef$Age_interval))-1

cor.test(amostra_Coef$estimate, amostra_Coef$Age_intervalnumber)

cor.test(filter(amostra_Coef, Age_intervalnumber > 1)$estimate, filter(amostra_Coef, Age_intervalnumber > 1)$Age_intervalnumber)

fig_slope_ageinterval <- ggplot(amostra_Coef, aes(y = estimate, x = Age_interval)) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
    geom_point() +
  geom_line(group = 1) + 
  geom_hline(yintercept = 1.25, linetype = "dashed") +
    theme_pubr() + 
    theme(axis.text.x = element_text(angle = 45, hjust = 1, size=9)
    ) + labs(y ="Slope (after harmonization)", x = "Age interval") +  geom_text(aes(label = N), nudge_y = 0.2)

fig_slope_ageinterval

# ggplot(filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL", Age_interval != "[0,5)", Age_interval != "[95,100]"), aes(y = K_shiftc, x = Age_interval), aes(y = K, x = Age_interval)) +
#     geom_boxplot() + 
#     theme_pubr() + 
#     theme(axis.text.x = element_text(angle = 45, hjust = 1, size=9)
#     ) + labs(y ="K (after harmonization)", x = "Age interval")

# ggsave("fig_slope_ageinterval.png", plot = fig_slope_ageinterval, dpi=1200, width = 18, height = 14, units = "cm", device = "png")
ggsave("fig_slope_ageinterval.pdf", plot = fig_slope_ageinterval, dpi=1200, width = 18, height = 14, units = "cm", device = "pdf")
```


```{r eval=FALSE, include=FALSE}
amostra_Coef <- filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL" | Diagnostic == "AD" , Age_interval != "[0,5)", Age_interval != "[95,100]") %>%
    group_by(Age_interval, Diagnostic) %>%
    do(fit_amostra = tidy(lm(1/2 * log10(AvgThickness_shiftc) + log10(TotalArea_shiftc) ~ log10(ExposedArea_shiftc), data = ., na.action = na.omit), conf.int = TRUE, conf.level = 0.95)) %>% 
    unnest(fit_amostra)

amostras_Coef_age <- filter(dados_datasetscomp, ROI == "hemisphere", Diagnostic == "CTL"| Diagnostic == "AD" , Age_interval != "[0,5)", Age_interval != "[95,100]") %>%
    group_by(Age_interval, Diagnostic) %>%
    summarise(
        N = n_distinct(SUBJ),
        min_age = min(Age),
        max_age = max(Age))

amostra_Coef <- full_join(amostra_Coef, amostras_Coef_age) %>% filter(term == "log10(ExposedArea_shiftc)")

amostra_Coef$Diagnostic <- factor(amostra_Coef$Diagnostic, levels=c('AD', 'CTL'))

ggplot(amostra_Coef, aes(y = estimate, x = Age_interval, color = Diagnostic)) +
    geom_errorbar(aes(ymin = conf.low, ymax = conf.high), width = 0.2) +
    geom_point() + geom_hline(yintercept = 1.25, linetype = "dashed") +
    theme_pubr() + 
    theme(axis.title = element_text(size = 11),
          axis.text = element_text(size = 10),
          text = element_text(size = 10),
          axis.text.x = element_text(angle = 45, hjust = 1, size=9)
    ) + labs(y ="Slope (after harmonization)", x = "Age interval") +  geom_text(aes(label = N), nudge_y = 0.3)

amostra_Coef%>% dplyr::select(-c(term)) %>% kable() %>% kable_styling()

```

## GM volume \~ Age

```{r cache=TRUE}
m.1 <- lme4::lmer(AvgThickness_shiftc ~ Age * Diagnostic +(1|Sample) + (1|Sample:Diagnostic) , data = filter(dados_datasetscomp, ROI == "hemisphere"))

# performance::check_model(m.1)

anova(m.1)
sigma.hat(m.1)
r.squaredGLMM(m.1)

mean <- lsmeans(m.1, ~ Diagnostic, var="Age")
m.lstk <- lstrends(m.1, ~ Diagnostic, var="Age")
pairs <- pairs(m.lstk) 

mean <- as_tibble(mean)
m.lstk <- as_tibble(m.lstk)

 lst <- m.lstk %>% mutate(
     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)))

tableT <- full_join(mean, lst, by = c("Diagnostic")) %>% mutate(percent = Age.trend*100/abs(lsmean), Variable = "T_shiftc")
 
# m.lstk %>% mutate(
#     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)),
#     CI = paste(signif(asymp.LCL, 2), ";", signif(asymp.UCL, 2)),
#   ) %>% dplyr::select(c(Diagnostic, TX, CI)) %>% full_join(N_SUBJ)

m.lstk %>% kable() %>% kable_styling()

coefs <- data.frame(coef(summary(m.1)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs

pairs %>% kable() %>% kable_styling()
 
# ggplot(data = as_tibble(pairs), aes(
#     x = contrast,
#     y = estimate,
#     ymin = estimate - SE,
#     ymax = estimate + SE)) +
#     geom_hline(yintercept = 0,
#                linetype = "11",
#                colour = "grey60") +
#     geom_pointrange( position = position_dodge(width = 0.2)) +
#    geom_text(aes(label = str_c("p adj ", signif(p.value, digits = 2))), nudge_x = 0.3, nudge_y = 0.0003) +
#     coord_flip() + 
#     labs(y =  expression('Differences in slopes ('*Delta*'AvgT/'*Delta*'Age)'), x = "Contrast") +
#   theme_pubr() + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10), text = element_text(size = 10))

# e <- effect("Age:Diagnostic", m.1)
# e <- as.data.frame(e)

# ggplot(e, aes(x = Age, y = fit,ymin=lower, ymax=upper, shape = Diagnostic)) + geom_pointrange() + geom_line() + theme_pubr() + labs(x = "Age [years]", y = "Avg Cortical Thickness", shape = "Diagnostic") + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10), text = element_text(size = 10))

# re <- as_tibble(ranef(m.1))

# ggplot(filter(re, grpvar == "Sample"), aes(y = condval, x = as.character(grp))) +
#   geom_errorbar(aes(ymin = condval+1.96*condsd, ymax = condval-1.96*condsd), width = 0.2) +
#   geom_point() + geom_hline(yintercept = 0, linetype = "dashed") +
#   theme_pubr() + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10),
#     text = element_text(size = 10),
#     axis.text.x = element_text(angle = 45, hjust = 1, size=9)
#   ) + labs(y ="T variation", x = "Sample")

```

## AvgThickness \~ Age

```{r cache=TRUE}
m.1 <- lme4::lmer(AvgThickness_shiftc ~ Age * Diagnostic +(1|Sample) + (1|Sample:Diagnostic) , data = filter(dados_datasetscomp, ROI == "hemisphere"))

# performance::check_model(m.1)

anova(m.1)
sigma.hat(m.1)
r.squaredGLMM(m.1)

mean <- lsmeans(m.1, ~ Diagnostic, var="Age")
m.lstk <- lstrends(m.1, ~ Diagnostic, var="Age")
pairs <- pairs(m.lstk) 

mean <- as_tibble(mean)
m.lstk <- as_tibble(m.lstk)

 lst <- m.lstk %>% mutate(
     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)))

tableT <- full_join(mean, lst, by = c("Diagnostic")) %>% mutate(percent = Age.trend*100/abs(lsmean), Variable = "T_shiftc")
 
# m.lstk %>% mutate(
#     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)),
#     CI = paste(signif(asymp.LCL, 2), ";", signif(asymp.UCL, 2)),
#   ) %>% dplyr::select(c(Diagnostic, TX, CI)) %>% full_join(N_SUBJ)

m.lstk %>% kable() %>% kable_styling()

coefs <- data.frame(coef(summary(m.1)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs

pairs %>% kable() %>% kable_styling()
 
# ggplot(data = as_tibble(pairs), aes(
#     x = contrast,
#     y = estimate,
#     ymin = estimate - SE,
#     ymax = estimate + SE)) +
#     geom_hline(yintercept = 0,
#                linetype = "11",
#                colour = "grey60") +
#     geom_pointrange( position = position_dodge(width = 0.2)) +
#    geom_text(aes(label = str_c("p adj ", signif(p.value, digits = 2))), nudge_x = 0.3, nudge_y = 0.0003) +
#     coord_flip() + 
#     labs(y =  expression('Differences in slopes ('*Delta*'AvgT/'*Delta*'Age)'), x = "Contrast") +
#   theme_pubr() + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10), text = element_text(size = 10))

# e <- effect("Age:Diagnostic", m.1)
# e <- as.data.frame(e)

# ggplot(e, aes(x = Age, y = fit,ymin=lower, ymax=upper, shape = Diagnostic)) + geom_pointrange() + geom_line() + theme_pubr() + labs(x = "Age [years]", y = "Avg Cortical Thickness", shape = "Diagnostic") + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10), text = element_text(size = 10))

# re <- as_tibble(ranef(m.1))

# ggplot(filter(re, grpvar == "Sample"), aes(y = condval, x = as.character(grp))) +
#   geom_errorbar(aes(ymin = condval+1.96*condsd, ymax = condval-1.96*condsd), width = 0.2) +
#   geom_point() + geom_hline(yintercept = 0, linetype = "dashed") +
#   theme_pubr() + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10),
#     text = element_text(size = 10),
#     axis.text.x = element_text(angle = 45, hjust = 1, size=9)
#   ) + labs(y ="T variation", x = "Sample")

```

## TotalArea \~ Age

```{r cache=TRUE}
m.1 <- lme4::lmer(TotalArea_shiftc ~ Age * Diagnostic +(1|Sample) + (1|Sample:Diagnostic) , data = filter(dados_datasetscomp, ROI == "hemisphere"))

# performance::check_model(m.1)

anova(m.1)
sigma.hat(m.1)
r.squaredGLMM(m.1)

mean <- lsmeans(m.1, ~ Diagnostic, var="Age")
m.lstk <- lstrends(m.1, ~ Diagnostic, var="Age")
pairs <- pairs(m.lstk) 

mean <- as_tibble(mean)
m.lstk <- as_tibble(m.lstk)

 lst <- m.lstk %>% mutate(
     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)))

tableAT <- full_join(mean, lst, by = c("Diagnostic")) %>% mutate(percent = Age.trend*100/abs(lsmean), Variable = "AT_shiftc")
 
# m.lstk %>% mutate(
#     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)),
#     CI = paste(signif(asymp.LCL, 2), ";", signif(asymp.UCL, 2)),
#   ) %>% dplyr::select(c(Diagnostic, TX, CI)) %>% full_join(N_SUBJ)

m.lstk %>% kable() %>% kable_styling()

coefs <- data.frame(coef(summary(m.1)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs

pairs %>% kable() %>% kable_styling()
 
# ggplot(data = as_tibble(pairs), aes(
#     x = contrast,
#     y = estimate,
#     ymin = estimate - SE,
#     ymax = estimate + SE)) +
#     geom_hline(yintercept = 0,
#                linetype = "11",
#                colour = "grey60") +
#     geom_pointrange( position = position_dodge(width = 0.2)) +
#    geom_text(aes(label = str_c("p adj ", signif(p.value, digits = 2))), nudge_x = 0.3, nudge_y = 0.0003) +
#     coord_flip() + 
#     labs(y =  expression('Differences in slopes ('*Delta*'AvgT/'*Delta*'Age)'), x = "Contrast") +
#   theme_pubr() + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10), text = element_text(size = 10))

# e <- effect("Age:Diagnostic", m.1)
# e <- as.data.frame(e)

# ggplot(e, aes(x = Age, y = fit,ymin=lower, ymax=upper, shape = Diagnostic)) + geom_pointrange() + geom_line() + theme_pubr() + labs(x = "Age [years]", y = "Avg Cortical Thickness", shape = "Diagnostic") + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10), text = element_text(size = 10))

# re <- as_tibble(ranef(m.1))

# ggplot(filter(re, grpvar == "Sample"), aes(y = condval, x = as.character(grp))) +
#   geom_errorbar(aes(ymin = condval+1.96*condsd, ymax = condval-1.96*condsd), width = 0.2) +
#   geom_point() + geom_hline(yintercept = 0, linetype = "dashed") +
#   theme_pubr() + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10),
#     text = element_text(size = 10),
#     axis.text.x = element_text(angle = 45, hjust = 1, size=9)
#   ) + labs(y ="T variation", x = "Sample")

```

## ExposedArea \~ Age

```{r cache=TRUE}
m.1 <- lme4::lmer(ExposedArea_shiftc ~ Age * Diagnostic +(1|Sample) + (1|Sample:Diagnostic) , data = filter(dados_datasetscomp, ROI == "hemisphere"))

# performance::check_model(m.1)

anova(m.1)
sigma.hat(m.1)
r.squaredGLMM(m.1)

mean <- lsmeans(m.1, ~ Diagnostic, var="Age")
m.lstk <- lstrends(m.1, ~ Diagnostic, var="Age")
pairs <- pairs(m.lstk) 

mean <- as_tibble(mean)
m.lstk <- as_tibble(m.lstk)

 lst <- m.lstk %>% mutate(
     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)))

tableAE <- full_join(mean, lst, by = c("Diagnostic")) %>% mutate(percent = Age.trend*100/abs(lsmean), Variable = "AE_shiftc")
 
# m.lstk %>% mutate(
#     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)),
#     CI = paste(signif(asymp.LCL, 2), ";", signif(asymp.UCL, 2)),
#   ) %>% dplyr::select(c(Diagnostic, TX, CI)) %>% full_join(N_SUBJ)

m.lstk %>% kable() %>% kable_styling()

coefs <- data.frame(coef(summary(m.1)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs

pairs %>% kable() %>% kable_styling()
 
# ggplot(data = as_tibble(pairs), aes(
#     x = contrast,
#     y = estimate,
#     ymin = estimate - SE,
#     ymax = estimate + SE)) +
#     geom_hline(yintercept = 0,
#                linetype = "11",
#                colour = "grey60") +
#     geom_pointrange( position = position_dodge(width = 0.2)) +
#    geom_text(aes(label = str_c("p adj ", signif(p.value, digits = 2))), nudge_x = 0.3, nudge_y = 0.0003) +
#     coord_flip() + 
#     labs(y =  expression('Differences in slopes ('*Delta*'AvgT/'*Delta*'Age)'), x = "Contrast") +
#   theme_pubr() + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10), text = element_text(size = 10))

# e <- effect("Age:Diagnostic", m.1)
# e <- as.data.frame(e)

# ggplot(e, aes(x = Age, y = fit,ymin=lower, ymax=upper, shape = Diagnostic)) + geom_pointrange() + geom_line() + theme_pubr() + labs(x = "Age [years]", y = "Avg Cortical Thickness", shape = "Diagnostic") + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10), text = element_text(size = 10))

# re <- as_tibble(ranef(m.1))

# ggplot(filter(re, grpvar == "Sample"), aes(y = condval, x = as.character(grp))) +
#   geom_errorbar(aes(ymin = condval+1.96*condsd, ymax = condval-1.96*condsd), width = 0.2) +
#   geom_point() + geom_hline(yintercept = 0, linetype = "dashed") +
#   theme_pubr() + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10),
#     text = element_text(size = 10),
#     axis.text.x = element_text(angle = 45, hjust = 1, size=9)
#   ) + labs(y ="T variation", x = "Sample")

```

## GI \~ Age

```{r cache=TRUE}
m.1 <- lme4::lmer(localGI_shiftc ~ Age * Diagnostic +(1|Sample) + (1|Sample:Diagnostic) , data = filter(dados_datasetscomp, ROI == "hemisphere"))

# performance::check_model(m.1)

anova(m.1)
sigma.hat(m.1)
r.squaredGLMM(m.1)

mean <- lsmeans(m.1, ~ Diagnostic, var="Age")
m.lstk <- lstrends(m.1, ~ Diagnostic, var="Age")
pairs <- pairs(m.lstk) 

mean <- as_tibble(mean)
m.lstk <- as_tibble(m.lstk)

lst <- m.lstk %>% mutate(
    TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)))

tableGI <- full_join(mean, lst, by = c("Diagnostic")) %>% mutate(percent = Age.trend*100/abs(lsmean), Variable = "localGI_shiftc")

# m.lstk %>% mutate(
#     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)),
#     CI = paste(signif(asymp.LCL, 2), ";", signif(asymp.UCL, 2)),
#   ) %>% dplyr::select(c(Diagnostic, TX, CI)) %>% full_join(N_SUBJ)
# 
# m.lstk %>% kable() %>% kable_styling()

coefs <- data.frame(coef(summary(m.1)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs

pairs %>% kable() %>% kable_styling()
 
# ggplot(data = as_tibble(pairs), aes(
#     x = contrast,
#     y = estimate,
#     ymin = estimate - SE,
#     ymax = estimate + SE)) +
#     geom_hline(yintercept = 0,
#                linetype = "11",
#                colour = "grey60") +
#     geom_pointrange( position = position_dodge(width = 0.2)) +
#    geom_text(aes(label = str_c("p adj ", signif(p.value, digits = 2))), nudge_x = 0.3, nudge_y = 0.0003) +
#     coord_flip() + 
#     labs(y =  expression('Differences in slopes ('*Delta*'GI/'*Delta*'Age)'), x = "Contrast") +
#   theme_pubr() + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10), text = element_text(size = 10))

# e <- effect("Age:Diagnostic", m.1)
# e <- as.data.frame(e)

# ggplot(e, aes(x = Age, y = fit,ymin=lower, ymax=upper, shape = Diagnostic)) + geom_pointrange() + geom_line() + theme_pubr() + labs(x = "Age [years]", y = "GI", shape = "Diagnostic") + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10), text = element_text(size = 10))

```

## K \~ Age

```{r cache=TRUE}
m.1 <- lme4::lmer(K_shiftc ~ Age * Diagnostic +(1|Sample) + (1|Sample:Diagnostic) , data = filter(dados_datasetscomp, ROI == "hemisphere"))

# performance::check_model(m.1)

anova(m.1)
sigma.hat(m.1)
r.squaredGLMM(m.1)

mean <- lsmeans(m.1, ~ Diagnostic, var="Age")
m.lstk <- lstrends(m.1, ~ Diagnostic, var="Age")
pairs <- pairs(m.lstk) 

mean <- as_tibble(mean)
m.lstk <- as_tibble(m.lstk)

lst <- m.lstk %>% mutate(
    TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)))

tableK <- full_join(mean, lst, by = c("Diagnostic")) %>% mutate(percent = Age.trend*100/abs(lsmean), Variable = "K_shiftc")

# m.lstk %>% mutate(
#     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)),
#     CI = paste(signif(asymp.LCL, 2), ";", signif(asymp.UCL, 2)),
#   ) %>% dplyr::select(c(Diagnostic, TX, CI)) %>% full_join(N_SUBJ) 
# m.lstk %>% kable() %>% kable_styling()

coefs <- data.frame(coef(summary(m.1)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs

pairs %>% kable() %>% kable_styling()
 
fig_TX_K_age_c <- ggplot(data = as_tibble(pairs), aes(
    x = contrast,
    y = estimate,
    ymin = estimate - SE,
    ymax = estimate + SE)) +
    geom_hline(yintercept = 0,
               linetype = "11",
               colour = "grey60") +
    geom_pointrange( position = position_dodge(width = 0.2)) +
   geom_text(aes(label = str_c("p adj ", signif(p.value, digits = 2))), nudge_x = 0.3, nudge_y = 0.0003) +
    coord_flip() + 
    labs(y =  expression('Differences in slopes ('*Delta*'K/'*Delta*'Age)'), x = "Contrast") +
  theme_pubr() + 
  theme(axis.title = element_text(size = 11),
    axis.text = element_text(size = 10), text = element_text(size = 10))

e <- effect("Age:Diagnostic", m.1)
e <- as.data.frame(e)

fig_K_age_c <- ggplot(e, aes(x = Age, y = fit,ymin=lower, ymax=upper, shape = Diagnostic)) + geom_pointrange() + geom_line() + theme_pubr() + labs(x = "Age [years]", y = "K", shape = "Diagnostic") + 
  theme(axis.title = element_text(size = 11),
    axis.text = element_text(size = 10), text = element_text(size = 10))

# re <- as_tibble(ranef(m.1))

# ggplot(filter(re, grpvar == "Sample"), aes(y = condval, x = as.character(grp))) +
#   geom_errorbar(aes(ymin = condval+1.96*condsd, ymax = condval-1.96*condsd), width = 0.2) +
#   geom_point() + geom_hline(yintercept = 0, linetype = "dashed") +
#   theme_pubr() + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10),
#     text = element_text(size = 10),
#     axis.text.x = element_text(angle = 45, hjust = 1, size=9)
#   ) + labs(y ="K variation", x = "Sample")

```


## S \~ Age

```{r cache=TRUE}
m.1 <- lme4::lmer(S_shiftc ~ Age * Diagnostic +(1|Sample) + (1|Sample:Diagnostic) , data = filter(dados_datasetscomp, ROI == "hemisphere"))

# performance::check_model(m.1)

anova(m.1)
sigma.hat(m.1)
r.squaredGLMM(m.1)

mean <- lsmeans(m.1, ~ Diagnostic, var="Age")
m.lstk <- lstrends(m.1, ~ Diagnostic, var="Age")
pairs <- pairs(m.lstk) 

mean <- as_tibble(mean)
m.lstk <- as_tibble(m.lstk)

lst <- m.lstk %>% mutate(
    TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)))

tableS <- full_join(mean, lst, by = c("Diagnostic")) %>% mutate(percent = Age.trend*100/abs(lsmean), Variable = "S_shiftc")

# m.lstk %>% mutate(
#     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)),
#     CI = paste(signif(asymp.LCL, 2), ";", signif(asymp.UCL, 2)),
#   ) %>% dplyr::select(c(Diagnostic, TX, CI)) %>% full_join(N_SUBJ)
# m.lstk %>% kable() %>% kable_styling()

coefs <- data.frame(coef(summary(m.1)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs

pairs %>% kable() %>% kable_styling()
 
fig_TX_S_age_c <- ggplot(data = as_tibble(pairs), aes(
    x = contrast,
    y = estimate,
    ymin = estimate - SE,
    ymax = estimate + SE)) +
    geom_hline(yintercept = 0,
               linetype = "11",
               colour = "grey60") +
    geom_pointrange( position = position_dodge(width = 0.2)) +
   geom_text(aes(label = str_c("p adj ", signif(p.value, digits = 2))), nudge_x = 0.3, nudge_y = 0.0003) +
    coord_flip() + 
    labs(y =  expression('Differences in slopes ('*Delta*'S/'*Delta*'Age)'), x = "Contrast") +
  theme_pubr() + 
  theme(axis.title = element_text(size = 11),
    axis.text = element_text(size = 10), text = element_text(size = 10))

e <- effect("Age:Diagnostic", m.1)
e <- as.data.frame(e)

fig_S_age_c <- ggplot(e, aes(x = Age, y = fit,ymin=lower, ymax=upper, shape = Diagnostic)) + geom_pointrange() + geom_line() + theme_pubr() + labs(x = "Age [years]", y = "S", shape = "Diagnostic") + 
  theme(axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    text = element_text(size = 10), legend.position = "none")

# re <- as_tibble(ranef(m.1)) 

# ggplot( filter(re, grpvar == "Sample"), aes(y = condval, x = as.character(grp))) +
#   geom_errorbar(aes(ymin = condval+1.96*condsd, ymax = condval-1.96*condsd), width = 0.2) +
#   geom_point() + geom_hline(yintercept = 0, linetype = "dashed") +
#   theme_pubr() + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10),
#     text = element_text(size = 10),
#     axis.text.x = element_text(angle = 45, hjust = 1, size=9)
#   ) + labs(y = "S variation", x = "Sample")

```

## I \~ Age

```{r cache=TRUE}
m.1 <- lme4::lmer(I_shiftc ~ Age * Diagnostic +(1|Sample) + (1|Sample:Diagnostic) , data = filter(dados_datasetscomp, ROI == "hemisphere"))

# performance::check_model(m.1)

anova(m.1)
sigma.hat(m.1)
r.squaredGLMM(m.1)

mean <- lsmeans(m.1, ~ Diagnostic, var="Age")
m.lstk <- lstrends(m.1, ~ Diagnostic, var="Age")
pairs <- pairs(m.lstk) 

mean <- as_tibble(mean)
m.lstk <- as_tibble(m.lstk)

lst <- m.lstk %>% mutate(
    TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)))

tableI <- full_join(mean, lst, by = c("Diagnostic")) %>% mutate(percent = Age.trend*100/abs(lsmean), Variable = "I_shiftc")

# m.lstk %>% mutate(
#     TX = paste(signif(Age.trend, 2), "±", signif(SE, 2)),
#     CI = paste(signif(asymp.LCL, 2), ";", signif(asymp.UCL, 2)),
#   ) %>% dplyr::select(c(Diagnostic, TX, CI)) %>% full_join(N_SUBJ)
# m.lstk %>% kable() %>% kable_styling()

coefs <- data.frame(coef(summary(m.1)))
coefs$p.z <- 2 * (1 - pnorm(abs(coefs$t.value)))
coefs

pairs %>% kable() %>% kable_styling()
 
fig_TX_I_age_c <- ggplot(data = as_tibble(pairs), aes(
    x = contrast,
    y = estimate,
    ymin = estimate - SE,
    ymax = estimate + SE)) +
    geom_hline(yintercept = 0,
               linetype = "11",
               colour = "grey60") +
    geom_pointrange( position = position_dodge(width = 0.2)) +
   geom_text(aes(label = str_c("p adj ", signif(p.value, digits = 2))), nudge_x = 0.3, nudge_y = 0.0003) +
    coord_flip() + 
    labs(y =  expression('Differences in slopes ('*Delta*'I/'*Delta*'Age)'), x = "Contrast") +
  theme_pubr() + 
  theme(axis.title = element_text(size = 11),
    axis.text = element_text(size = 10), text = element_text(size = 10))

e <- effect("Age:Diagnostic", m.1)
e <- as.data.frame(e)

fig_I_age_c <- ggplot(e, aes(x = Age, y = fit,ymin=lower, ymax=upper, shape = Diagnostic)) + geom_pointrange() + geom_line() + theme_pubr() + labs(x = "Age [years]", y = "I", shape = "Diagnostic") + 
  theme(axis.title = element_text(size = 11),
    axis.text = element_text(size = 10),
    text = element_text(size = 10), legend.position = "none")

# re <- as_tibble(ranef(m.1))

# ggplot(filter(re, grpvar == "Sample"), aes(y = condval, x = as.character(grp))) +
#   geom_errorbar(aes(ymin = condval+1.96*condsd, ymax = condval-1.96*condsd), width = 0.2) +
#   geom_point() + geom_hline(yintercept = 0, linetype = "dashed") +
#   theme_pubr() + 
#   theme(axis.title = element_text(size = 11),
#     axis.text = element_text(size = 10),
#     text = element_text(size = 10),
#     axis.text.x = element_text(angle = 45, hjust = 1, size=9)
#   ) + labs(y = "I variation", x = "Sample")

```

## Figures

```{r eval=FALSE, include=FALSE}
fig_TX_age_c <- ggarrange(fig_TX_K_age_c, ggarrange(fig_TX_S_age_c, fig_TX_I_age_c,labels = c("B","C"), nrow = 1, ncol = 2, font.label = list(size = 11)),labels = c("A"), nrow = 2, ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")

# fig_TX_age_c

fig_age_c <- ggarrange(fig_K_age_c, ggarrange(fig_S_age_c, fig_I_age_c,labels = c("B","C"), nrow = 1, ncol = 2, font.label = list(size = 11)),labels = c("A"), nrow = 2, ncol = 1, font.label = list(size = 11), common.legend = TRUE, legend = "top")

# fig_age_c

ggsave("fig_TX_age_c.pdf", plot = fig_TX_age_c, dpi=1200, width = 18, height = 18, units = "cm", device = "pdf")
# ggsave("fig_TX_age_c.png", plot = fig_TX_age_c, dpi=1200, width = 18, height = 18, units = "cm", device = "png")

ggsave("fig_age_c.pdf", plot = fig_age_c, dpi=1200, width = 18, height = 18, units = "cm", device = "pdf")
# ggsave("fig_age_c.png", plot = fig_age_c, dpi=1200, width = 18, height = 18, units = "cm", device = "png")

```

## rates
```{r echo=FALSE}
tableT <- tableT %>% dplyr::select(c(Diagnostic, percent, Variable, Age.trend, SE.y))
tableAT <- tableAT %>% dplyr::select(c(Diagnostic, percent, Variable, Age.trend, SE.y))
tableAE <- tableAE %>% dplyr::select(c(Diagnostic, percent, Variable, Age.trend, SE.y))
tableK <- tableK %>% dplyr::select(c(Diagnostic, percent, Variable, Age.trend, SE.y))
tableS <- tableS %>% dplyr::select(c(Diagnostic, percent, Variable, Age.trend, SE.y))
tableI <- tableI %>% dplyr::select(c(Diagnostic, percent, Variable, Age.trend, SE.y))
tableGI <- tableGI %>% dplyr::select(c(Diagnostic, percent, Variable, Age.trend, SE.y))

rate <- full_join(tableT, tableAT) %>%
  full_join(tableAE) %>%
  full_join(tableK) %>%
  full_join(tableS) %>%
  full_join(tableI) %>%
  full_join(tableGI) %>%
mutate(TX = paste(signif(Age.trend, 2), "±", signif(SE.y, 2), "(",signif(percent, 2), ")")) %>%
  dplyr::select(c(Variable, Diagnostic, TX))

rate %>%
  kable() %>%
  kable_styling()


write.csv(rate, "rate.csv")
```

